---
title: "Air Quality Warning" 
type: "community_smoke"
date: "`r Sys.Date()`"
params: 
  ice: Continue
  date: "`r Sys.Date()`"
  pollutant: PM10
  burnRestrictions: "Yes - Ben"
  sel_aqMet: "Sakshi"
  station: "Prince George"
  issuedate: "2025-07-20"
  HAauth: ""
  outputFormat: ""
  customText: "Current conditions are expected to persist for the next 24 hours."
format:
  markdown: default
  pdf:
    date-format: long
    include-in-header: 
      - src/header.tex
    mainfont: BCSans-Regular
    mainfontoptions:
      - Boldfont = BCSans-Bold
      - ItalicFont = BCSans-Italic
    fontsize: "11pt"
    urlcolor: bcblue
---
```{=html}
<!--
Copyright 2025 Province of British Columbia

This work is licensed under the Creative Commons Attribution 4.0 International License.
To view a copy of this license, visit http://creativecommons.org/licenses/by/4.0/.
-->
```

```{r setup, echo=FALSE, include=FALSE, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(echo=FALSE,
                      message=FALSE,
                      warning=FALSE)

library(tidyverse)
library(knitr)
library(data.table)
library(shiny)
library(sf)
library(lubridate)
library(readr)
library(xtable)

source(here::here("load_metadata.r"))
```

```{r health_tips, include=FALSE}

current_month <- as.numeric(format(Sys.Date(), "%m"))

health_tips <- paste(
  "# Actions you can take\n\n",
  
  "As air contaminant levels increase, health risks increase. Consider reducing or rescheduling outdoor sports, activities and events.\n\n",
  "People more likely to be negatively impacted by outdoor air pollution should reduce or reschedule strenuous activities outdoors or seek medical attention if experiencing symptoms. This includes people aged 65 and older, pregnant individuals, infants and young children, people with an existing illness or chronic health condition such as chronic obstructive pulmonary disease (COPD), heart disease and diabetes, and people who work outdoors.\n\n",
  
  "## Follow your common sense\n",
  "- Stop or reduce your activity level if breathing becomes uncomfortable or you feel unwell.\n",
  if (current_month %in% 6:8) {
    "- During warm weather, stay cool and drink plenty of fluids.\n"
  } else {
    ""
  },
  "- Always carry any rescue medications with you.\n",
  "- Make sure that children and others who cannot care for themselves follow the same advice.\n\n",
  
  "## Monitor your symptoms\n",
  "- Different people have different responses to elevated levels of air contaminants.\n",
  "- Mild irritation and discomfort such as eye, nose and throat irritation, headaches or a mild cough are common, and usually disappear when the air contaminants return to typical levels.\n",
  "- More serious but less common symptoms include wheezing, chest pains or severe cough.\n",
  "- People with asthma or other chronic illness should follow any personal care plans designed with their family physicians.\n",
  "- If you are unsure whether you need medical care, call HealthLink BC at 8-1-1.\n",
  "- If you are experiencing difficulty in breathing, chest pain or discomfort, or a severe cough, contact your physician, walk-in clinic, or emergency department.\n",
  "- If you are having a medical emergency, call 9-1-1.\n\n",
  
  "## Tips to reduce your exposure to air contaminants\n",
  "- When indoors, keep windows and doors closed as much as possible.\n",
  "- Air contaminant levels may be lower indoors but will still be elevated, so stay aware of your symptoms even when you are indoors.\n",
  if (current_month %in% 6:8) {
    "- When there is an extreme heat event occurring with poor air quality, prioritize keeping cool.\n"
  } else {
    ""
  },
  if (params$pollutant %in% c("PM25", "PM10")) {
    "- Protect your indoor air from outdoor air pollution. Actions can include using a clean, good quality air filter in your ventilation system and/or a certified portable air cleaner that can filter fine particles. Do-it yourself air cleaners may also be used if other options are not available. For details, visit: https://rb.gy/epi7qv.\n"
    "- If you must spend time outdoors, a well-constructed, well-fitting and properly worn respirator type mask (such as a NIOSH-certified N95 or equivalent respirator) can reduce your exposure to particulate matter. Even though exposure may be reduced, there can still be risks to health.\n"
  } else {
    ""
  },
  if ((params$pollutant == "PM25" && current_month %in% 6:8)) {
  "- Avoid roads with heavy vehicle traffic.\n"
  } else {
  ""
  },
  
  if (params$pollutant %in% c("PM25", "PM10") && current_month %in% c(9:12, 1:5)) {
    "- Avoid roads with heavy vehicle traffic and areas with wood smoke.\n"
  } else {
    ""
  },
  sep = ""
)

```

```{r additional_info, include=FALSE}

additional_info_PM25 <- paste0(
    "Fine particulate matter refers to airborne solid or liquid droplets with diameters of 2.5 micrometers ($\\mu$m) or less.  PM$_{2.5}$ levels tend to be highest around busy roads, industrial operations and neighbourhoods with residential wood burning. PM$_{2.5}$ can easily penetrate indoors because of their small size. Common sources of PM$_{2.5}$ that contribute to episodes of poor air quality  vary seasonally but can include wood smoke (from wood stoves and/or open burning) as well as emissions from industry and transportation sources such as automobiles, trucks and rail traffic.
    
This episode is expected to continue until there is a change in the current weather system.
    
The provincial air quality objective for PM$_{2.5}$ is 25 micrograms per cubic metre ($\\mu$g/m$^{3}$) averaged over 24 hours. 24-hour average PM$_{2.5}$ concentrations are summarized below for ", params$station, " and nearest monitored communities at ", format(floor_date(Sys.time(), "hour"), "%H:%M"), " local time today:"
  )

additional_info_PM10 <- paste0(
  "Coarse particulate matter refers to airborne solid or liquid droplets with diameters between 2.5 and 10 micrometers ($\\mu$m). Together with fine particulate matter (airborne solid or liquid droplets with diameters of 2.5 $\\mu$m or less), these particles are referred to as PM$_{10}$. Sources of PM$_{10}$ contributing to this air quality episode include road dust from the emission of winter traction material along busy and dry road surfaces. PM$_{10}$ can easily penetrate indoors because of their small size.
  
  The current dusty conditions are caused by road traffic stirring up winter traction materials that have accumulated on roadways over the past winter.\n\n",
  if (params$station == "Prince George") {paste0(
  "For information on the City of Prince George’s Clean Air Bylaw (8266), please contact City Hall at 250-561-7600 or visit the City’s website at: [https://www.princegeorge.ca](https://www.princegeorge.ca/).\n\n")
  } else {NULL},
  "The provincial air quality objective for PM$_{10}$ is 50 micrograms per cubic metre ($\\mu$g/m$^{3}$), averaged over 24 hours. 24-hour average PM$_{10}$ concentrations are summarized below for ", params$station, " and nearest monitored communities at ", format(floor_date(Sys.time(), "hour"), "%H:%M"), " PST today:"
  )

additional_info_O3 <- paste0(
  "Ground-level ozone is formed when chemicals (such as nitrogen oxides and volatile organic compounds) react in the air when there is sunlight. Nitrogen oxides are emitted from vehicles, boilers, building heating, and other combustion processes. VOCs are emitted from burning fossil fuels, evaporation of solvents (including paint, varnishes and thinners), refining and storing fuel and agricultural activities. VOCs also come from natural sources such as vegetation.
  
The highest concentrations of ground-level ozone usually happen between mid-afternoon and early evening on summer days.
  
The provincial air quality objectives for ozone is an eight-hour objective of 62 parts per billion (ppb) and a one-hour average of 82 ppb for air quality advisories. Air quality advisories for ground-level ozone may be issued when current concentrations exceed or are expected to exceed either of these objectives. \n",
  "- The 8-hour average ozone concentration and 1-hour maximum concentration (within the past 8 hours) are summarized below for ", params$station, " and nearest monitored communities at ", format(floor_date(Sys.time(), "hour"), "%H:%M"), " PST today:"
  )
```

```{r volunary_reduction, include=FALSE}

voluntary_reduction_PM25 <- paste0(
   if (current_month %in% c(10:12, 1:4)) {
    "- Avoid using wood stoves and fireplaces unless it is the sole heating source. If wood burning is the sole heating source, burn dry, seasoned wood and ensure an adequate supply of combustion air.\n"
  } else {
    ""
  },
 "- Reduce vehicle use where possible and avoid idling vehicles.\n",
    sep=""
)

voluntary_reduction_PM10 <- paste0(
  "- Avoid driving on the road shoulder or other areas where road traction material has accumulated.\n",
  "-  When cleaning driveways and parking lots, lightly wet the area before sweeping. Avoid using leaf blowers to clean up dirt during spring clean up.\n"
)

voluntary_reduction_O3 <- paste0(
  "Fuel your vehicle in the cooler evening hours or after dark.\n",
  "Reduce or reschedule using other gasoline and diesel equipment, such as lawn mowers, trimmers and leaf blowers.\n"
)
```

```{r HAcontact, include=FALSE}

HAauth <- Health_Authority %>%
  filter(Station == params$station) %>%
  pull(Health_authority)

HAcontact <- health_contact |>
    filter(authority == HAauth) |>
    select(authority, contact) |>
    group_by(authority) |>
    summarise(html_string = paste(contact, collapse = "<br />"))

```

```{r pollutant_specific_clauses, include=FALSE}
if (params$pollutant == "PM25") {
  pollutantExpansion <- "fine particulate matter"
  additional_info <- additional_info_PM25
  voluntary_reduction <- voluntary_reduction_PM25
  if (params$ice == "Issue") {
  first_paragraph <- paste0("The Ministry of Environment and Parks in collaboration with the ", HAauth, " has issued an Air Quality Warning for ", params$station, " due to elevated ", pollutantExpansion, ".")
  } else if (params$ice == "Continue") {
  first_paragraph <- paste0("The Ministry of Environment and Parks in collaboration with the ", HAauth, " issued an Air Quality Warning for ", params$station, " on ", format(as.Date(params$issuedate), "%B %d, %Y"), ". The Air Quality Advisory remains in effect due to high concentrations of ", pollutantExpansion, " that are expected to persist until weather conditions change."
    )
}
} else if (params$pollutant == "PM10") {  
  pollutantExpansion <- "coarse particulate matter"
  additional_info <- additional_info_PM10
  voluntary_reduction <- voluntary_reduction_PM10
  if (params$ice == "Issue") {
  first_paragraph <- paste0("The Ministry of Environment and Parks in collaboration with the ", HAauth, " has issued an Air Quality Warning for ", params$station, " due to elevated ", pollutantExpansion, ".")
  } else if (params$ice == "Continue") {
  first_paragraph <- paste0("The Ministry of Environment and Parks in collaboration with the ", HAauth, " issued an Air Quality Warning for ", params$station, " on ", format(as.Date(params$issuedate), "%B %d, %Y"), ". The advisory remains in effect due to high concentrations of coarse particulate matter that are expected to persist until there is precipitation, dust suppression or a change in traffic patterns. Levels tend to be highest around busy roads and industrial operations. This advisory is in effect until further notice."
    )
}
} else if (params$pollutant == "O3") {
  pollutantExpansion <- "ground level ozone"
  additional_info <- additional_info_O3
  voluntary_reduction <- voluntary_reduction_O3
  first_paragraph <- paste0("The Ministry of Environment and Parks in collaboration with the ", HAauth, " has issued an Air Quality Warning for ", params$station, " due to elevated ", pollutantExpansion, ".")
} else {
  title_pollutant <- NULL
  background <- NULL
}
```

```{r title_ICE, include=FALSE}

title_ICE <- if (params$ice == "Issue") {
  "issued"
} else if (params$ice == "Continue") {
  "continued"
} else if (params$ice == "End") {
  "ended" 
} else {
  NULL
}

```

```{r PG_text, include=FALSE}

PG_text <- if (params$station == "Prince George") 
  {"The City of Prince George's Clean Air Bylaw prohibits all open burning, including backyard burning, and land clearing burning. During air quality advisories, this bylaw also prohibits recreational fires, use of wood-burning appliances (except for sole wood burning heat users), and street sweeping activities (unless approved by an Authorized Person)."
} else {
  NULL
}
```

```{r bylaws, include=FALSE}

bylaws <- if (params$station == "Burns Lake") {
  "Burns Lake bylaw 871 restricts the use of wood burning appliances during air quality advisories:\n\n
  - **6a**: No person shall use a Wood Burning Appliance at any time when an air quality advisory is in effect, except to heat the premises that are equipped with no heating appliance or facilities other than the Wood Burning Appliance.\n\n
  - Contact the Burns Lake municipal office at [250-692-7587](tel:2506927587) for more information on woodstove restrictions."
} else if (params$station == "Duncan") {
  "Comply with the requirements of local woodstove bylaws that apply to periods when air quality advisories are in effect (e.g., City of Duncan and Municipality of North Cowichan)."
} else if (params$station == "Houston") {
  "Houston bylaw 947 restricts the use of wood burning appliances during air quality advisories:\n\n
  - **2.1.1**: No person shall use a wood burning appliance at any time when an air quality advisory is in effect, except to heat premises that are equipped with no heating appliance or facilities other than the wood burning appliance.\n\n
  - Contact the Houston municipal office at [250-845-2238](tel:2508452238) for more information on woodstove restrictions."
} else if (params$station == "Prince George") {
  "The City of Prince George's **Clean Air Bylaw** prohibits all open burning, including backyard burning, and land clearing burning. During air quality advisories, this bylaw also prohibits **recreational fires**, use of **wood-burning appliances** (except for sole wood-burning heat users), and **street sweeping activities** (unless approved by an Authorized Person)."
} else if (params$station == "Smithers") {
  "Smithers bylaw 1520 restricts the use of wood burning appliances during air quality advisories:\n\n
  - **6a**: No person shall use a wood-burning appliance at any time when an air quality advisory is in effect, except to heat premises that are equipped with no heating appliance or facilities other than the wood burning appliance.\n\n
  - Contact the Smithers municipal office at [250-847-1600](tel:2508471600) for more information on woodstove restrictions."
} else if (params$station == "Valemount") {
  "Valemount Bylaw No. 838 restricts the use of wood burning appliances during air quality advisories:\n\n
  - **5.1**: A person will not use a wood-burning appliance at any time when an air quality advisory is in effect, except to heat premises that are equipped with no heating source other than the wood burning appliance.\n\n
  - Contact the Valemount municipal office at [250-566-4435](tel:2505664435) or visit [https://valemount.ca](https://valemount.ca) for more information on wood stove restrictions."
} else {
  NULL
}

# Only run if 'bylaws' is defined and a character string
if (exists("bylaws") && is.character(bylaws)) {

  # Split into lines
  bylaw_lines <- unlist(strsplit(bylaws, "\n"))

  # Clean up leading/trailing whitespace
  bylaw_lines <- trimws(bylaw_lines)

  # Remove empty lines
  bylaw_lines <- bylaw_lines[bylaw_lines != ""]

  # Separate bulleted (items) vs non-bulleted (intro) points
  bylaw_items <- bylaw_lines[grepl("^-", bylaw_lines)]
  bylaw_intro <- bylaw_lines[!grepl("^-", bylaw_lines)]

  # Create HTML
  formatted_bylaws <- paste0(
    if (length(bylaw_intro) > 0) paste0("<p>", paste(bylaw_intro, collapse = " "), "</p>") else NULL,
    if (length(bylaw_items) > 0) {
      paste0(
        "<ul>",
        paste0("<li>", sub("^-\\s*", "", bylaw_items), "</li>", collapse = ""),
        "</ul>"
      )
    } else {
      NULL
    }
  )

} else {
  # Fallback if there are no local bylaws
  formatted_bylaws <- NULL
}


```

```{r ban_areas, include=FALSE}
burn_areas <- c(
  "Burns Lake" = "Lakes Timber Supply Area",
  "Golden" = "15km of the Hospital",
  "Grand Forks" = "a 15 km radius of City Hall",
  "Houston" = "the Morice Timber Supply Area",
  "Powell River" = "Sunshine Coast Timber Supply Area",
  "Prince George" = "within the Prince George and Quesnel Natural Resource Districts in all locations east of Highway 97, south of Highway 16, and west of Bowron Lake Provincial Park",
  "Squamish" = "a 20 km radius of Squamish Municipal Hall",
  "Terrace" = "Kalum Resource District",
  "Vanderhoof" = "former Vanderhoof Resource District",
  "Whistler" = "Soo Timber Supply Area"
)

# Use single bracket to get a named element safely
ban_area <- burn_areas[params$station]

# If the station isn't found, return the default
ban_area <- ifelse(is.na(ban_area), "a 10 km radius of City Hall", ban_area)
```

```{r burn_ban_clauses, results='asis', echo=FALSE}

nextBulletin <- Sys.Date() + 1

if (params$burnRestrictions %in% c("Yes - Arvind", "Yes - Ben")) {
  
  title_burn_advisory <- "and open burning restrictions"
  mandatory_reductions_heading <- "# Mandatory Emission Reduction Actions"
  burn_ban <- paste0(
    "Open burning restrictions are now in effect for ", params$station, 
    " and surrounding area until ", format(as.Date(nextBulletin), "%B %d, %Y"), 
    " at 4:00 PM local time. No new fires may be initiated and no additional material may be added to existing fires. ",
    "For more information on burning restrictions, see the section below entitled **Mandatory Emission Reduction Actions**."
  )
  signature_style <- "margin-top:15px; margin-left: 20px;"  # with indent
  date_div <- paste0("<div style='margin-top:15px; margin-left: 20px;'>",
                       "<p>Date issued: ", format(Sys.Date(), "%B %d, %Y"), "</p>",
                       "</div>")
  
  # Define signature and title info depending on who it is
  if (params$burnRestrictions == "Yes - Arvind") {
    signature_img <- file.path("utility", "signatures", "SignatureAS.png")
    name_line <- "Arvind Saraswat"
    title_line1 <- "Director, Environmental Management Act"
    title_line2 <- "Environmental Monitoring and Analysis Branch"
  } else if (params$burnRestrictions == "Yes - Ben") {
    signature_img <- file.path("utility", "signatures", "SignatureBW.jpg")
    name_line <- "Benjamin Weinstein"
    title_line1 <- "For Director, Environmental Management Act"
    title_line2 <- "Environmental Monitoring and Analysis Branch"
    
  }
  
  burn_ban_bylaws <- paste0(
    "<ul>",
      "<li>As pollution is occurring or is likely to occur from open burning, pursuant to Sections 30(1) and 30(2) of the Open Burning Smoke Control Regulation, the Director has prohibited open burning in the ", 
      params$station, " area within ", ban_area, 
      " until ", format(as.Date(nextBulletin), "%B %d, %Y"), 
      " at 04:00 PM. No vegetative debris may be ignited or added to ignited piles. Contravention of these provisions may be subject to a fine under the Regulation.</li>",
    "</ul>",
    date_div,
    "<div style='", signature_style, "'>",
      "<img src='", signature_img, "' style='height:60px;'><br>",
      name_line, "<br>",
      title_line1, "<br>",
      title_line2,
    "</div>",
    "<ul style='margin-top:20px;'>",
     if (!is.null(formatted_bylaws) && nzchar(formatted_bylaws)) {
      paste0(
        "<li>Additional local restrictions may apply under municipal bylaws:",
        formatted_bylaws,
        "</li>"
      )
    } else {
      NULL
    },
    "<li>Industry is required to follow permit requirements that are triggered during air quality advisories and are asked to reduce emissions wherever possible.</li>",
  "</ul>"
)
  
} else {
  
  title_burn_advisory <- NULL
  mandatory_reductions_heading <- NULL
  burn_ban <- NULL
  burn_ban_bylaws <- NULL
  bylaws <- NULL
}

```

```{r table-setup, include=FALSE}

pollutant <- params$pollutant
stations <- params$station

COMPLETE_THRESHOLD_PCT <- 75

# ---- Check that stations are valid for this pollutant ----
valid_stations <- buddy_stations %>%
  filter(Station %in% stations) %>%
  pull(Buddy_location) %>%
  unique()

if (length(valid_stations) == 0) {
  stop("None of the specified stations are valid for this pollutant.")
}

# ---- Lookup table for units and averaging times ----
lookup <- data.frame(
  units = c("$\\mu$g/m$^{3}$", "$\\mu$g/m$^{3}$", "ppb"),
  hours = c(24, 24, 8),
  row.names = c("PM25", "PM10", "O3")
) %>%
  mutate(avgtimelabel = sprintf("%d-hr", hours))

# ---- Read in CSV from FTP ----
columndefs <- cols_only(
  DATE_PST = col_character(),
  STATION_NAME = col_factor(),
  RAW_VALUE = col_double(),
  REPORTED_VALUE = col_double(),
  INSTRUMENT = col_factor(),
  UNITS = col_character(),
  PARAMETER = col_character(),
  EMS_ID = col_character()
)

dfraw <- read_csv(
  sprintf("ftp://ftp.env.gov.bc.ca/pub/outgoing/AIR/Hourly_raw_air_data/Air_Quality/%s.csv", pollutant),
  col_types = columndefs
)

# ---- Clean and filter data ----
df <- dfraw %>%
  mutate(DATE_PST = as.POSIXct(DATE_PST, format = "%Y-%m-%d %H:%M", tz = "Etc/GMT+8")) %>%
  filter(INSTRUMENT != "") %>%
  filter(STATION_NAME %in% valid_stations)


# ---- Get time window per station ----
dts <- df %>%
  group_by(STATION_NAME) %>%
  summarize(max_dt = max(DATE_PST), .groups = "drop") %>%
  mutate(min_dt = max_dt - (3600 * (lookup[pollutant, "hours"] - 1)))

df_filtered <- df %>%
  left_join(dts, by = "STATION_NAME") %>%
  filter(DATE_PST >= min_dt & DATE_PST <= max_dt) %>%
  select(-min_dt, -max_dt)

# ---- Summarize concentrations ----
integrated <- df_filtered %>%
  group_by(STATION_NAME, PARAMETER) %>%
  summarize(
    n_hours = n(),
    n_na = sum(is.na(REPORTED_VALUE)),
    pct_valid = 100 * (n_hours - n_na) / n_hours,
    mean_conc = if_else(n_hours == n_na, NA_real_, mean(REPORTED_VALUE, na.rm = TRUE)),
    max_conc = if_else(n_hours == n_na, NA_real_, max(REPORTED_VALUE, na.rm = TRUE)),
    start_dt = min(DATE_PST),
    end_dt = max(DATE_PST),
    .groups = "drop"
  ) %>%
  mutate(
    mean_conc = na_if(mean_conc, pct_valid < COMPLETE_THRESHOLD_PCT),
    max_conc = na_if(max_conc, pct_valid < COMPLETE_THRESHOLD_PCT)
  )

stationByParameterList <- read_rds(sprintf("./utility/stations_%s.rds", tolower(pollutant)))

# Join with station info using STATION_NAME
integrated <- integrated %>%
  left_join(
    stationByParameterList %>% select(STATION_NAME, COMMUNITY),
    by = "STATION_NAME"
  )

# ---- Format wide table ----
table2 <- integrated %>%
  mutate(
    mean_conc = ifelse(is.finite(mean_conc), sprintf("%.1f", mean_conc), "NA"),
    max_conc = ifelse(is.finite(max_conc), sprintf("%.1f", max_conc), "NA")
  ) %>%
  pivot_wider(
    names_from = PARAMETER,
    id_cols = c(STATION_NAME, COMMUNITY),
    values_from = c(mean_conc, max_conc)
  ) %>%
  rename(Station = STATION_NAME, Community = COMMUNITY)

# ---- Clean up table: remove suffixes ----
table2 <- table2 %>%
  rename_with(~ gsub(sprintf("_%s", pollutant), "", .x), everything())

table2 <- table2[ , -1]

# ---- Transpose and finalize ----
pt2 <- t(table2) %>% as.data.frame() %>% rownames_to_column()
colnames(pt2) <- pt2[1, ]
pt3 <- pt2[-c(1), ]  # drop row names and station name if only one station

if (pollutant == "O3") {
    pt4 <- pt3
  } else {
    pt4 <- pt3 %>% filter(Community != "max_conc")
}

# ---- Final format ----
final_table <- pt4 %>%
  mutate(Community = case_when(
    Community == "mean_conc" ~ sprintf("%s average (%s)", lookup[pollutant, "avgtimelabel"], lookup[pollutant, "units"]),
    Community == "max_conc" ~ sprintf("Max. within %s (%s)", gsub("-", " ", lookup[pollutant, "avgtimelabel"]), lookup[pollutant, "units"]),
    TRUE ~ Community
  ))

```

```{r ENVcontact, include=FALSE}

ENVcontact <- aq_mets |>
  filter(nickname == params$sel_aqMet) |>
  mutate(contact = paste(fullname_typeset, title, ministry, phone, sep = "<br />")) |> 
  pull(contact)
```

```{r}
#| label: create-logo-list

logos_list <- list(
  "Government of British Columbia" = "BCID_V_RGB_pos",
  "First Nations Health Authority" = "FNHA",
  "Interior Health Authority" = "IH", 
  "Fraser Health Authority" = "FH", 
  "Vancouver Coastal Health Authority" = "VCH", 
  "Vancouver Island Health Authority" = "VIH", 
  "Northern Health Authority" = "NH")

# Logos selected by user and ordered as per logos_list
# Logos selected by user
logos_selected_names <- c("Government of British Columbia", HAauth)

# Get the correct logo codes (e.g., "BCID_V_RGB_pos", "VCH")
logos_selected <- logos_list[logos_selected_names]  # this returns a named list

# Build image lines
# logo_image_line <- c()
# for (logo_name in names(logos_selected)) {
#   logo_image_line <- c(
#     logo_image_line,
#     paste0("![", logo_name, " logo](/utility/logos/logo_", logos_selected[[logo_name]], ".png){width=150px}\\")
#   )
# }

n_logos <- length(logos_selected)

logo_image_line <- c()
logos_combined <- c()

# Set logo path based on output format
if (params$outputFormat == "markdown") {
  logo_path <- " logo](//assets/logo_"

  # Build a vector of quarto lines to insert image for each logo
  # It is more efficient to use `sapply` but this might be more readable

  # Add each logo's insert line to the vector
  for (logo_name in names(logos_selected)) {
    logo_image_line <- c(logo_image_line,
                         paste0("![", logo_name, logo_path, logos_selected[[logo_name]], ".png)\\"))
  }}

if (params$outputFormat == "pdf") {
  
  logo_path <- "https://github.com/bcgov/aqwarnings/blob/main/frontend/assets/logo_"
  logo_urls <- c() #start with empty vector
  
  # create vector of urls
  for (logo_name in names(logos_selected)) {
    logo_urls <- c(
      logo_urls, 
      paste0(logo_path,
             logos_selected[[logo_name]],
             ".png?raw=true"
             )
    )
  }

# format logos
logos <- magick::image_read(logo_urls) |> 
  magick::image_trim() |> # remove all white space
  magick::image_border(color = "none", geometry = "100x25") |> # add uniform white space to top and right side of each image
  magick::image_scale("x200") # scale

# Combine logos into single image and scale
logos_combined <- magick::image_append(logos) |> 
  magick::image_write(path = "logo.png")
}

```

<!-- the trailing slash means the text in square brackets is alt text -->

[`r paste("::: {layout-ncol=", n_logos," layout-valign=\"bottom\"}")`]{.content-visible when-format="markdown"}


```{r}
#| label: print-logos
#| results: asis

# `cat` is used to avoid extra processing
# `sep` argument adds the line break and then a new line as required
if (params$outputFormat == "markdown") cat(logo_image_line, sep="\n\n") 
```

[`r paste(":::")`]{.content-visible when-format="markdown"}
# Air quality warning `r title_burn_advisory` `r title_ICE` for `r params$station` - `r format(Sys.Date(), "%B %d, %Y")`

`r first_paragraph`

Exposure to `r pollutantExpansion` is particularly a concern for individuals with chronic conditions such as asthma, COPD, heart disease, and diabetes; respiratory infections such as COVID-19, pregnant women, infants, and older adults. Persons with chronic underlying medical conditions or acute infections should postpone or reduce strenuous exercise until the advisory is lifted. Where appropriate, maintain physical distancing. Anyone experiencing symptoms such as continuing eye or throat irritation, chest discomfort, shortness of breath, cough or wheezing, should follow the advice of their health care provider. Staying indoors helps to reduce exposure.

`r burn_ban`

`r PG_text`

Real-time air quality observations and information regarding the health effects of air pollution can be found at: [Air Quality Data](https://www.gov.bc.ca/airquality).

`r params$customText`

The next update will be on `r format(as.Date(nextBulletin), "%B %d, %Y")` and posted to the province’s [Air Quality Warnings webpage](https://www.gov.bc.ca/airquality).  

`r health_tips`

`r mandatory_reductions_heading`

```{r, echo=FALSE, results='asis'}
knitr::asis_output(burn_ban_bylaws)
```

# Voluntary Emission Reductions

`r voluntary_reduction`

# Additional Information

`r additional_info`

```{r final-table}
kable(final_table, row.names = FALSE)
```

## For general information about air quality in British Columbia: 

- Real-time air quality observations and information regarding the health effects of air pollution can be found at [gov.bc.ca/airquality](https://gov.bc.ca/airquality).
- View the latest [Air Quality Health Index (AQHI) and forecast][What's the air like today?]
- Visit the provincial [air quality map][Latest air quality data] to view real-time air quality measurements
- [Metro Vancouver’s Air Quality Data and Warnings][Air quality data and warnings]

[Air quality data and warnings]: https://metrovancouver.org/services/air-quality-climate-action/air-quality-data-and-advisories
[What's the air like today?]: https://www.env.gov.bc.ca/epd/bcairquality/data/aqhi-table.html
[Latest air quality data]: https://www.env.gov.bc.ca/epd/bcairquality/readings/find-stations-map.html

# Contact Information

[{{< card_start width="wide" >}}]{.content-visible when-format="markdown"}`r ENVcontact`[{{< card_end >}}]{.content-visible when-format="markdown"}

[{{< card_start width="wide" >}}]{.content-visible when-format="markdown"}`r HAcontact$html_string`[{{< card_end >}}]{.content-visible when-format="markdown"}

