---
date: "`r Sys.Date()`"
params:
  ice: "Issue" # Issue; Continue
  pollutant: "PM25 & PM10"
  aqMet: "Sakshi Jain"
  location: "Prince George"
  issuedate: "2025-07-20"
  nextUpdate: "2025-09-28"
  outputFormat: "markdown"
  customMessage: ""
  burnRestrictions: 0   # 0 = No; 1 = Yes (Ben); 2 = Yes (Arvind)
  burnRestrictionArea: ""
  burnRestrictionEndDate: "2025-09-28"
  burnRestrictionEndTime: "12:00 PM"
  warningLevel: "yellow"
format:
  markdown: default
  pdf:
    mainfont: "Roboto"
    date-format: long
    include-in-header:
      - src/header.tex
    fontsize: "12pt"
    urlcolor: bcblue
---

```{=html}
<!--
Copyright 2025 Province of British Columbia

This work is licensed under the Creative Commons Attribution 4.0 International License.
To view a copy of this license, visit http://creativecommons.org/licenses/by/4.0/.
-->
```

```{r setup, echo=FALSE, include=FALSE, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(echo=FALSE,
                      message=FALSE,
                      warning=FALSE)

library(tidyverse)   #could this be streamlined?
library(knitr)
library(data.table)
library(shiny)
library(sf)
library(lubridate)
library(readr)
library(xtable)
library(tibble)
library(dplyr)

source(here::here("load_metadata.r"))
source(here::here(file.path("functions", "local_tz.r")))
source(here::here(file.path("src", "construct_banners.r")))
source(here::here(file.path("src", "construct_logos.r")))
source(here::here(file.path("src", "construct_pollutants_table.r")))
source(here::here(file.path("src", "local_emissions_bylaws.r")))
source(here::here(file.path("src", "construct_shared_dependencies.r")))

# date/time information
numeric_current_month <- as.numeric(format(Sys.Date(), "%m"))
# create single line break based on output file format
if (params$outputFormat == "markdown") markup_separator <- "<br />" else markup_separator <- "  \n"

run_pg <- (params$location == "Prince George")  #to control conditional text related to Prince George only
```

```{r table-setup, include=FALSE}

data_pollutants <- strsplit(params$pollutant, "&")[[1]] |> trimws()
min_data_capture_pct <- 75

local_results <- construct_pollutants_table(data_pollutants, params$location, min_data_capture_pct)
string_local_hour <- local_results$local_hour
data_pollutants_table <- local_results$data_pollutants_table

# Expose additional results as data_pollutants_table1, data_pollutants_table2, etc.
list2env(local_results$results, envir = .GlobalEnv)
```

```{r health-authority, include=FALSE}
#| results: asis
health_authority_details <- get_health_authority_details(params$location, markup_separator)
string_health_authority <- health_authority_details$health_authority
data_health_authority_contact <- health_authority_details$health_authority_contact
```
```{r voluntary_reduction, include=FALSE}

if (numeric_current_month %in% c(1:4, 10:12)) {
  string_voluntary_reduction_PM25 <- paste0(
    "- Avoid using wood stoves and fireplaces unless it is the sole heating source. If wood burning is the sole heating source, burn dry, seasoned wood and ensure an adequate supply of combustion air.\n",
    "- Reduce vehicle use where possible and avoid idling vehicles.\n",
    "- Compost or mulch garden waste intead of burning it.\n",
    "- Take public transit or carpool whenever possible."
    )
} else {
  paste0(
    "- Reduce vehicle use where possible and avoid idling vehicles.\n",
    "- Compost or mulch garden waste intead of burning it.\n",
    "- Take public transit or carpool whenever possible."
    )
  }

string_voluntary_reduction_PM10 <- paste0(
  "- Avoid driving on the road shoulder or other areas where road traction material has accumulated.\n",
  "- When cleaning driveways and parking lots, lightly wet the area before sweeping\n",
  "- Avoid using leaf blowers to clean up dirt during spring clean up.\n"
  )

string_voluntary_reduction_PM <- paste0(
  string_voluntary_reduction_PM25,
  string_voluntary_reduction_PM10,
  sep = "\n\n"
)

string_voluntary_reduction_O3 <- paste0(
  "- Fuel your vehicle in the cooler evening hours or after dark.\n",
  "- Reduce or reschedule using other gasoline and diesel equipment, such as lawn mowers, trimmers and leaf blowers.\n"
)
```

```{r pollutant_specific_clauses, include=FALSE}

# Lookup table for pollutants
pollutant_tbl <- tribble(
  ~pollutant,~pollutantExpansion,~voluntary_reduction,
  "PM25","fine particulate matter",string_voluntary_reduction_PM25,
  "PM10","coarse particulate matter",string_voluntary_reduction_PM10,
  "PM25 & PM10","fine particulate matter and coarse particulate matter",string_voluntary_reduction_PM,
  "O3","ground level ozone",string_voluntary_reduction_O3
)

# Pick the row for the current pollutant
pollutant_info <- pollutant_tbl |> filter(pollutant == params$pollutant)

# Assign values
string_pollutant_expansion  <- pollutant_info$pollutantExpansion
string_voluntary_reduction <- pollutant_info$voluntary_reduction

# Look up table for warning level dynamic text in intro and actions you can take sections
tbl_level <- tribble(
  ~level,
  ~intro_description,
  ~intro_action,
  ~intro_indoors,
  ~action_first_paragraph,
  "yellow", 
  "elevated ", 
  "Postpone or reduce ", 
  "Spend time ",
  "Exposure to poor air quality can affect your health. Reduce the duration and intensity of outdoor activities. Spend time indoors in a space with filtered air.",
  "orange", 
  "very high levels of ", 
  "Avoid ",
  "Stay ",
  "Exposure to high levels of air contaminants pose a threat to your health regardless of age or health status. Avoid outdoor activities. Stay indoors in a space with filtered air."
)

# Pick the row for the current warning level
level_text <- tbl_level |> filter(level == params$warningLevel)

# Assign values
string_level_intro_description <- level_text$intro_description
string_level_intro_action <- level_text$intro_action
string_level_intro_indoors <- level_text$intro_indoors
string_level_action_first_paragraph <- level_text$action_first_paragraph

# Build first_paragraph
if (params$ice == "Issue") {
  first_paragraph <- paste0(
    "The Ministry of Environment and Parks in collaboration with the ",
    string_health_authority, " has issued an Air Quality Warning for ", params$location,
    " due to ", string_level_intro_description, string_pollutant_expansion, "."
  )
} else if (params$ice == "Continue") {
  first_paragraph <- paste0(
    "The Ministry of Environment and Parks in collaboration with the ",
    string_health_authority, " issued an Air Quality Warning for ", params$location,
    " on ", format(as.Date(params$issuedate), "%B %d, %Y"),
    ". The warning remains in effect due to elevated ", string_pollutant_expansion, "."
  )
}

```

```{r more_info, include=FALSE, results='asis'}

reference_more_info_table <- tibble(
  pollutant = c("PM25", "PM10", "O3"),
  info = c(
    paste0(
      "Fine particulate matter refers to airborne solid or liquid droplets with diameters of 2.5 micrometers (μm) or less.  PM~2.5~ levels tend to be highest around busy roads, industrial operations and neighbourhoods with residential wood burning. PM~2.5~ can easily penetrate indoors because of their small size. Common sources of PM~2.5~ that contribute to episodes of poor air quality  vary seasonally but can include wood smoke (from wood stoves and/or open burning) as well as emissions from industry and transportation sources such as automobiles, trucks and rail traffic.

The provincial air quality objective for PM~2.5~ is 25 micrograms per cubic metre (μg/m^3^) averaged over 24 hours. 24-hour average PM~2.5~ concentrations are summarized below for ",
params$location,
" and nearest monitored communities at ",
string_local_hour,
" local time today:"
    ),
paste0(
  "Coarse particulate matter refers to airborne solid or liquid droplets with diameters between 2.5 and 10 micrometers (μm). Together with fine particulate matter (airborne solid or liquid droplets with diameters of 2.5 μm or less), these particles are referred to as PM~10~. Sources of PM~10~ contributing to this air quality episode include road dust from the emission of winter traction material along busy and dry road surfaces. PM~10~ can easily penetrate indoors because of their small size.

  The current dusty conditions are caused by road traffic stirring up winter traction materials that have accumulated on roadways over the past winter.

  The provincial air quality objective for PM~10~ is 50 micrograms per cubic metre (μg/m^3^), averaged over 24 hours. 24-hour average PM~10~ concentrations are summarized below for ",
  params$location,
  " and nearest monitored communities at ",
  string_local_hour,
  " local time today:"
),
paste0(
  "Ground-level ozone is formed when chemicals (such as nitrogen oxides and volatile organic compounds) react in the air when there is sunlight. Nitrogen oxides are emitted from vehicles, boilers, building heating, and other combustion processes. VOCs are emitted from burning fossil fuels, evaporation of solvents (including paint, varnishes and thinners), refining and storing fuel and agricultural activities. VOCs also come from natural sources such as vegetation.

The highest concentrations of ground-level ozone usually happen between mid-afternoon and early evening on summer days.

The provincial air quality objectives for ozone is an eight-hour objective of 62 parts per billion (ppb) and a one-hour average of 82 ppb for air quality warnings. Air quality warnings for ground-level ozone may be issued when current concentrations exceed or are expected to exceed either of these objectives.

The 8-hour average ozone concentration and 1-hour maximum concentration (within the past 8 hours) are summarized below for ",
params$location,
" and nearest monitored communities at ",
string_local_hour,
" local time today:"
)
  )
)

local_results_info <- list()

for (i in seq_along(data_pollutants)) {

  string_more_pollutant_info <- reference_more_info_table |>
    filter(pollutant == data_pollutants[i]) |>
    pull(info)

    local_results_info[[paste0("string_more_pollutant_info", i)]] <- string_more_pollutant_info
}

# Expose them as string_more_pollutant_info1, string_more_pollutant_info2, etc.
list2env(local_results_info, envir = .GlobalEnv)
```

```{r bylaws, include=FALSE}

string_bylaws <- build_bylaw_information(params$location)
```

```{r burn_ban, results='asis', echo=FALSE}

if (params$burnRestrictions > 0) {

  # Define statutory decision maker (sdm)
  if (params$burnRestrictions == 1) {
    name_line <- "Benjamin Weinstein"
    title_line1 <- "For Director, Environmental Management Act"
    title_line2 <- "Environmental Monitoring and Analysis Branch"

  } else if (params$burnRestrictions == 2) {
   name_line <- "TBD"
   title_line1 <- "Director, Environmental Management Act"
   title_line2 <- "Environmental Monitoring and Analysis Branch"
  }

  # to do: improve formatting - single line spacing for sdm information, \n and <br> don't seem to work; rather than using paste could this text be added to a conditional code block directly in the document?
  string_burn_ban <- paste0(
  "- NOTICE of DIRECTOR under s.30 Open Burning Smoke Control Regulation:\n\n",
  "> As pollution is occurring or is likely to occur from open burning, pursuant to Sections 30(1) and 30(2) of the Open Burning Smoke Control Regulation, the Director has prohibited open burning within ",
  params$burnRestrictionArea,
  " until ",
  format(as.Date(params$burnRestrictionEndDate), "%B %d, %Y"),
  " ",
  params$burnRestrictionEndTime,
  " local time. No vegetative debris may be ignited or added to ignited piles. Contravention of these provisions may be subject to a fine under the Regulation.\n\n",
  "&emsp;&emsp;&nbsp;Date issued: ", format(Sys.Date(), "%B %d, %Y"), markup_separator,
  "&emsp;&emsp;&nbsp;", name_line, markup_separator,
  "&emsp;&emsp;&nbsp;", title_line1, markup_separator,
  "&emsp;&emsp;&nbsp;", title_line2
  )

} else {
  string_burn_ban <- NULL
}

```


```{r create-logo-list}
logos <- construct_logos(params$outputFormat, string_health_authority)
```

<!-- Logo header, the layout-col should be set based on number of logos including FHNA and BCGov -->

<!-- the trailing slash means the text in square brackets is alt text -->

[`r paste("::: {layout-ncol=", logos$n_logos," layout-valign=\"bottom\"}")`]{.content-visible when-format="markdown"}

```{r print-logos}
#| results: asis

# `cat` is used to avoid extra processing
# `sep` argument adds the line break and then a new line as required
if (params$outputFormat == "markdown") cat(logos$logo_image_line, sep="\n\n")
```

[`r paste(":::")`]{.content-visible when-format="markdown"}

`r build_banners(params$outputFormat, params$warningLevel == "yellow", params$warningLevel == "orange", params$warningLevel == "red")`

`r first_paragraph`

Exposure to `r string_pollutant_expansion` is particularly a concern for infants, older adults, individuals with chronic conditions (such as asthma, COPD, heart disease, and diabetes) or respiratory infections, and those who are pregnant.

`r string_level_intro_action` strenuous outdoor activities until the warning has ended. `r string_level_intro_indoors` indoors in a space with filtered air. Monitor your symptoms and seek prompt medical attention if needed.

```{r, results='asis', echo=FALSE}
if (params$burnRestrictions > 0) {
  cat(
    "Open burning restrictions are now in effect within ", params$burnRestrictionArea,
    ". No new fires may be initiated, and no additional material may be added to existing fires. For more information on burning restrictions, refer to the Mandatory Emission Reduction Actions section below.",
    sep = ""
  )
}
```

`r params$customMessage`

The next update will be on `r format(as.Date(params$nextUpdate), "%B %d, %Y")` and posted to the
province’s [Air Quality Warnings webpage](https://www.gov.bc.ca/airquality).

Visit the
provincial [air quality data webpage](https://www2.gov.bc.ca/gov/content/environment/air-land-water/air/air-quality) for
real-time observations.

## Actions you can take

`r string_level_action_first_paragraph`

Some people are more likely to be negatively impacted by outdoor air pollution. 
This includes people with lung and heart conditions, pregnant individuals, infants and young children, 
people aged 65 and older, and anyone whose health is affected by a chronic condition or an acute illness. 

::: {.content-visible when-format="markdown"}
{{< accordion_controls >}} {{< accordion_start title="Follow your common sense" initiallyOpen="true" >}}
:::

::: {.content-visible when-format="pdf"}
### Follow your common sense
:::
- Stop or reduce your activity level if breathing becomes uncomfortable or you feel unwell.
```{r, results='asis', echo=FALSE}

if (numeric_current_month %in% 6:8) {
  cat("- During warm weather, stay cool and drink plenty of fluids.\n")
}
```
- Always carry any rescue medications with you.
- Make sure that children and others who cannot care for themselves follow the same advice.

::: {.content-visible when-format="markdown"}
{{< accordion_end >}} {{< accordion_start title="Monitor your symptoms" initiallyOpen="true" >}}
:::

::: {.content-visible when-format="pdf"}
### Monitor your symptoms
:::

- Different people have different responses to elevated levels of air contaminants.
- Mild irritation and discomfort such as eye, nose and throat irritation, headaches or a mild cough are common, and
  usually disappear when the air contaminants return to typical levels.
- More serious but less common symptoms include wheezing, chest pains or severe cough.
- People with asthma or other chronic illness should follow any personal care plans designed with their family
  physicians.
- If you are unsure whether you need medical care, call HealthLink BC at 8-1-1.
- If you are experiencing difficulty in breathing, chest pain or discomfort, or a severe cough, contact your physician, 
  walk-in clinic, or emergency department. **If you are having a medical emergency, call 9-1-1.**

::: {.content-visible when-format="markdown"}
{{< accordion_end >}} {{< accordion_start title="Tips to reduce your exposure to air pollution" initiallyOpen="true" >}}
:::

::: {.content-visible when-format="pdf"}
### Tips to reduce your exposure to air pollution
:::

- Air contaminant levels may be elevated indoors, so stay aware of your symptoms even when you are indoors.

```{r, results='asis', echo=FALSE}

if (numeric_current_month %in% 6:8) {
  cat(
    "- When indoors, keep windows and doors closed.  If it is hot, prioritize keeping cool.\n",
    "- When there is an extreme heat event ocurring with poor air quality , prioritize keeping cool.\n",
    "- If your home is too hot, spend time in places with air conditioning such as community centres, libraries, or malls.\n"
    )
}

if (numeric_current_month %in% c(1:5, 9:12)) {
  cat("- When indoors, keep windows and doors closed.\n")
}

if (params$pollutant %in% c("PM25", "PM10", "PM25 & PM10")) {
  cat(
    "- Protect your indoor air from outdoor air pollution. Use a clean, good quality air filter in your ventilation system and/or a certified portable air cleaner that can filter fine particles. Avoid models that generate ozone, as these can produce harmful ozone concentrations indoors. Do-it-yourself air cleaners may also be used if other options are unavailable. For more details, see the BC Centre of Control [Do-it-yourself Air Cleaners](https://www.bccdc.ca/resource-gallery/Documents/Guidelines%20and%20Forms/Guidelines%20and%20Manuals/Health-Environment/BCCDC_WildFire_FactSheet_BoxFanAirFilters.pdf) fact sheet.\n", 
    "- If you must be outside, a well-constructed, well-fitting and properly worn respirator type mask (such as a NIOSH-certified N95 or equivalent respirator) can reduce your exposure to particulate matter. Even though exposure may be reduced, there can still be risks to health.\n",
    "- When in a vehicle, keep windows closed with air set to recirculate."
  )
}

if (params$pollutant == "O3") {
  cat("-Air cleaners can help improve indoor air quality, but avoid models that generate ozone, as these can produce harmful ozone concentrations indoors.")
}
```

::: {.content-visible when-format="markdown"}
{{< accordion_end >}}
:::

## Emission reduction actions

::: {.content-visible when-format="markdown"}
{{< accordion_start title="Mandatory actions" initiallyOpen="true" >}}
:::

::: {.content-visible when-format="pdf"}
### Mandatory actions
:::

- Facilities with air discharge authorizations under the Environmental Management Act are required to follow trigger
  actions within their permit related to Air Quality Warnings and are encouraged to reduce any other emissions where
  possible.

`r if (!is.null(string_bylaws) && nzchar(string_bylaws)) paste("-", string_bylaws)`
`r if (params$burnRestrictions > 0) paste(string_burn_ban)`

::: {.content-visible when-format="markdown"}
{{< accordion_end >}} {{< accordion_start title="Voluntary actions" initiallyOpen="true" >}}
:::

::: {.content-visible when-format="pdf"}
### Voluntary actions
:::

`r string_voluntary_reduction`

::: {.content-visible when-format="markdown"}
{{< accordion_end >}}
:::

## More information

::: {.content-visible when-format="markdown"}
{{< accordion_start title="Current conditions" initiallyOpen="true" >}}
:::

```{r pollutants-table1, results='asis', echo=FALSE}
if (length(data_pollutants) == 1) {
  cat(string_more_pollutant_info, "\n\n")
} else { # pollutants == c(PM2.5, PM10)
  cat(string_more_pollutant_info1, "\n\n")
}
cat('[::: {.table-responsive}]{.content-visible when-format="markdown"}\n')
if (length(data_pollutants) == 1) {
  print(knitr::kable(data_pollutants_table, row.names = FALSE, escape = FALSE))
  } else {  # pollutants == c(PM2.5, PM10)
    print(knitr::kable(data_pollutants_table1, row.names = FALSE))
  }
cat('[\n:::\n]{.content-visible when-format="markdown"}')
```

```{r pollutants-table2, results='asis', echo=FALSE, eval= (all(c("PM25", "PM10") %in% data_pollutants))}
cat("\n\n", string_more_pollutant_info2, "\n\n")
cat('[::: {.table-responsive}]{.content-visible when-format="markdown"}\n')
print(knitr::kable(data_pollutants_table2, row.names = FALSE, escape = FALSE))
cat('[\n:::]{.content-visible when-format="markdown"}\n')
```

::: {.content-visible when-format="markdown"}
{{< accordion_end >}}
:::

::: {.content-visible when-format="pdf"}
### For additional information about air quality in British Columbia
:::

::: {.content-visible when-format="markdown"}
**For additional information about air quality in British Columbia:**

  ::: bcds-card-wrapper
{{< card_start title="Air Quality Health Index" variant="info" logo="/assets/icon_air.svg" >}} Provincial summary of
latest index and forecast.

[What's the air like today?](https://www.env.gov.bc.ca/epd/bcairquality/data/aqhi-table.html) {{< card_end >}} {{<
card_start title="Provincal Air Quality Map" variant="info" logo="/assets/icon_air.svg" >}} Provincial air quality data
refreshed every hour.

[Latest air quality data](https://www.env.gov.bc.ca/epd/bcairquality/readings/find-stations-map.html) {{<
card_end >}} {{< card_start title="Metro Vancouver Air Quality" variant="success" logo="/assets/icon_chart.svg" >}}
Warnings and a map of air quality and weather data.

[Air quality data and warnings](https://metrovancouver.org/services/air-quality-climate-action/air-quality-data-and-advisories) {{<
card_end >}}
  ```{r, results='asis', echo=FALSE, eval= run_pg}

  cat('{{< card_start title="Local Resources" variant="info" logo="/assets/icon_air.svg" >}}\n\n [Prince George Air Improvement Roundtable](https://www.pgairquality.com/) {{< card_end >}}\n\n')
  ```
  :::
:::

::: {.content-visible when-format="pdf"}
- View the 
  latest [Air Quality Health Index (AQHI) and forecast](https://www.env.gov.bc.ca/epd/bcairquality/data/aqhi-table.html)
- Visit the provincial [air quality map](https://www.env.gov.bc.ca/epd/bcairquality/readings/find-stations-map.html) to 
  view real-time air quality measurements
- Visit the provincial [air quality data webpage](https://www2.gov.bc.ca/gov/content/environment/air-land-water/air/air-quality) for more information about air quality in B.C.
- Visit Metro 
  Vancouver's [air quality data and warnings page](https://metrovancouver.org/services/air-quality-climate-action/air-quality-data-and-advisories)
- Subscribe to receive air quality warnings on
  the [Air Quality Subscriptions Service](https://www2.gov.bc.ca/gov/content/environment/air-land-water/air/air-quality/air-advisories/air-quality-subscription-service) page
```{r, results='asis', echo=FALSE, eval= run_pg}
  cat("- Visit the [Prince George Air Improvement Roundtable's page](https://www.pgairquality.com/)")
```
:::

## Contact

::: {.content-visible when-format="pdf"}
### Media questions regarding this Air Quality Warning
:::

::: {.content-visible when-format="markdown"}
**Media questions regarding this Air Quality Warning:** {{< card_start width="wide" >}}
:::

`r get_meteorologist_contact(params$aqMet, markup_separator)`

::: {.content-visible when-format="markdown"}
{{< card_end >}}
:::

::: {.content-visible when-format="pdf"}
### Media questions regarding health implications when air quality is degraded
:::

::: {.content-visible when-format="markdown"}
**Media questions regarding health implications when air quality is degraded:** {{< card_start width="wide" >}}
:::

`r paste(data_health_authority_contact$html_string, collapse = "\n\n")`

::: {.content-visible when-format="markdown"}
{{< card_end >}}
:::
