---
date: "`r Sys.Date()`"
params:
  aqMet: "Sakshi Jain"
  nearestMonitor: "Whistler"
  issuedate: "2025-07-09"
  outputFormat: "markdown"
  customMessage: ""
  burnRestrictionArea: ""
format:
  markdown: default
  pdf:
    mainfont: "Roboto"
    date-format: long
    include-in-header:
      - src/header.tex
    fontsize: "12pt"
    urlcolor: bcblue
---

```{=html}
<!--
Copyright 2025 Province of British Columbia

This work is licensed under the Creative Commons Attribution 4.0 International License.
To view a copy of this license, visit http://creativecommons.org/licenses/by/4.0/.
-->
```

```{r setup, echo=FALSE, include=FALSE, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(echo=FALSE,
                      message=FALSE,
                      warning=FALSE)

library(tidyverse)
library(knitr)
library(data.table)
library(shiny)
library(sf)
library(lubridate)
library(readr)
library(xtable)

source(here::here("load_metadata.r"))
source(here::here(file.path("functions", "local_tz.r")))
source(here::here(file.path("src", "construct_logos.r")))
source(here::here(file.path("src", "construct_pollutants_table.r")))
source(here::here(file.path("src", "construct_shared_dependencies.r")))

# create single line break based on output file format
if (params$outputFormat == "markdown") sep_type <- "<br />" else sep_type <- "  \n"
```

```{r table-setup, include=FALSE}
pollutants <- "PM25"
min_data_capture_pct <- 75

results <- construct_pollutants_table(pollutants, params$nearestMonitor, min_data_capture_pct)
local_hour <- results$local_hour
final_table <- results$final_table
```


```{r health-authority, include=FALSE}
#| results: asis

health_authority_details <- get_health_authority_details(params$location)
HAauth <- health_authority_details$health_authority
HAcontact <- health_authority_details$health_authority_contact
```

```{r pollutant_specific_clauses, include=FALSE}
pollutant_tbl <- tribble(
  ~pollutant,~pollutantExpansion,
  "PM25","fine particulate matter")

# Pick the row for the current pollutant
pollutant_info <- pollutant_tbl |> filter(pollutant == "PM25")

# Assign values
pollutantExpansion  <- pollutant_info$pollutantExpansion

  first_paragraph <- paste0(
    "The Ministry of Environment and Parks in collaboration with the ", HAauth,
    " has ended the Pollution Prevention Notice and open burning restrictions within ",
    params$burnRestrictionArea,
    " that were issued on ",
    format(as.Date(params$issuedate), "%B %d, %Y"),
    "."
  )

```

```{r additional_info, include=FALSE}

additional_info_PM25 <- paste0(
    "The provincial air quality objective for PM~2.5~ is 25 micrograms per cubic metre (μg/m^3^) averaged over 24 hours. 24-hour average PM~2.5~ concentrations are summarized below for ", params$nearestMonitor, " and nearest monitored communities at ", local_hour, " local time today:"
  )

```

```{r create-logo-list}
logos <- construct_logos(params$outputFormat, HAauth)
```

<!-- Logo header, the layout-col should be set based on number of logos including FHNA and BCGov -->
<!-- the trailing slash means the text in square brackets is alt text -->

[`r paste("::: {layout-ncol=", logos$n_logos," layout-valign=\"bottom\"}")`]{.content-visible when-format="markdown"}

```{r print-logos}
#| results: asis

# `cat` is used to avoid extra processing
# `sep` argument adds the line break and then a new line as required
if (params$outputFormat == "markdown") cat(logos$logo_image_line, sep="\n\n")
```

[`r paste(":::")`]{.content-visible when-format="markdown"}

`r first_paragraph`

`r params$customMessage`

## More information

::: {.content-visible when-format="markdown"}
{{< accordion_start title="Current conditions" initiallyOpen="true" >}}
:::

::: {.content-visible when-format="pdf"}
### Current conditions
:::

```{r final-table, results='asis', echo=FALSE}

# Single pollutant: PM25
  cat(additional_info_PM25, "\n\n")
  cat('[::: {.table-responsive}]{.content-visible when-format="markdown"}\n')
  print(knitr::kable(final_table, row.names = FALSE, escape = FALSE))
  cat('[\n:::]{.content-visible when-format="markdown"}\n')
```

Visit the provincial [air quality data webpage](https://www2.gov.bc.ca/gov/content/environment/air-land-water/air/air-quality) for real-time observations.

::: {.content-visible when-format="markdown"}
{{< accordion_end >}}
:::

## Contact

::: {.content-visible when-format="pdf"}
### Media questions regarding this Pollution Prevention Notice
:::

::: {.content-visible when-format="markdown"}
**Media questions regarding this Pollution Prevention Notice:**
{{< card_start width="wide" >}}
:::

`r get_meteorologist_contact(params$aqMet, sep_type)`

::: {.content-visible when-format="markdown"}
{{< card_end >}}
:::

::: {.content-visible when-format="pdf"}
### Media questions regarding health implications when air quality is degraded
:::

::: {.content-visible when-format="markdown"}
**Media questions regarding health implications when air quality is degraded:**
{{< card_start  width="wide" >}}
:::

`r paste(HAcontact$html_string, collapse = "\n\n")`

::: {.content-visible when-format="markdown"}
{{< card_end >}}
:::
