---
date: "`r Sys.Date()`"
params:
  ice: "Issue" # Issue; Continue
  aqMet: "Sakshi Jain"
  nearestMonitor: "Prince George"
  issuedate: "2025-07-20"
  nextUpdate: "2025-09-28"
  outputFormat: "markdown"
  customMessage: ""
  burnRestrictions: 1   # 1 = Yes (Ben); 2 = Yes (TBD)
  burnRestrictionArea: ""
  burnRestrictionEndDate: "2025-09-28"
  burnRestrictionEndTime: "12:00 PM"
format:
  markdown: default
  pdf:
    mainfont: "Roboto"
    date-format: long
    include-in-header:
      - src/header.tex
    fontsize: "12pt"
    urlcolor: bcblue
---
```{=html}
<!--
Copyright 2025 Province of British Columbia

This work is licensed under the Creative Commons Attribution 4.0 International License.
To view a copy of this license, visit http://creativecommons.org/licenses/by/4.0/.
-->
```

```{r setup, echo=FALSE, include=FALSE, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(echo=FALSE,
                      message=FALSE,
                      warning=FALSE)

library(tidyverse)   #could this be streamlined?
library(knitr)
library(data.table)
library(shiny)
library(sf)
library(lubridate)
library(readr)
library(xtable)
library(tibble)
library(dplyr)

source(here::here("load_metadata.r"))
source(here::here(file.path("functions", "local_tz.r")))
source(here::here(file.path("src", "construct_logos.r")))
source(here::here(file.path("src", "construct_pollutants_table.r")))
source(here::here(file.path("src", "construct_shared_dependencies.r")))

# date/time information
current_month <- as.numeric(format(Sys.Date(), "%m"))
# create single line break based on output file format
if (params$outputFormat == "markdown") markup_separator <- "<br />" else markup_separator <- "  \n"
run_pg <- (params$nearestMonitor == "Prince George")  #to control conditional text related to Prince George only
```

```{r table-setup, include=FALSE}

data_pollutants <- "PM25"
min_data_capture_pct <- 75

local_results <- construct_pollutants_table(data_pollutants, params$nearestMonitor, min_data_capture_pct)
string_local_hour <- local_results$local_hour
data_pollutants_table <- local_results$data_pollutants_table
```

```{r health-authority, include=FALSE}
#| results: asis

health_authority_details <- get_health_authority_details(params$nearestMonitor, markup_separator)
string_health_authority <- health_authority_details$health_authority
data_health_authority_contact <- health_authority_details$health_authority_contact
```

```{r voluntary_reduction, include=FALSE}

if (current_month %in% c(1:4, 10:12)) {
  voluntary_reduction_PM25 <- paste0(
    "- Avoid using wood stoves and fireplaces unless it is the sole heating source. If wood burning is the sole heating source, burn dry, seasoned wood and ensure an adequate supply of combustion air.\n",
    "- Reduce vehicle use where possible and avoid idling vehicles.\n",
    "- Compost or mulch garden waste intead of burning it.\n",
    "- Take public transit or carpool whenever possible."
    )
} else {
  paste0(
    "- Reduce vehicle use where possible and avoid idling vehicles.\n",
    "- Compost or mulch garden waste intead of burning it.\n",
    "- Take public transit or carpool whenever possible."
    )
  }
```

```{r pollutant_specific_clauses, include=FALSE}

# Lookup table for pollutants
data_pollutant_tbl <- tribble(
  ~pollutant,~string_pollutant_expansion,~voluntary_reduction,
  "PM25","fine particulate matter",voluntary_reduction_PM25,
)

# Pick the row for the current pollutant
data_pollutant_info <- data_pollutant_tbl |> filter(pollutant == "PM25")

# Assign values
string_pollutant_expansion  <- data_pollutant_info$pollutantExpansion
voluntary_reduction <- data_pollutant_info$voluntary_reduction

# Build first_paragraph
if (params$ice == "Issue") {
  first_paragraph <- paste0(
    "The Ministry of Environment and Parks in collaboration with the ",
    string_health_authority, " has issued a Pollution Prevention Notice within ", params$burnRestrictionArea,
    " due to elevated ", string_pollutant_expansion, "."
  )
} else if (params$ice == "Continue") {
  first_paragraph <- paste0(
    "The Ministry of Environment and Parks in collaboration with the ",
    string_health_authority, " issued Pollution Prevention Notice within ", params$burnRestrictionArea,
    " on ", format(as.Date(params$issuedate), "%B %d, %Y"),
    ". The warning remains in effect due to elevated ", string_pollutant_expansion, "."
  )
}

```

```{r additional_info, include=FALSE, results='asis'}

additional_info_PM25 <- paste0("Fine particulate matter refers to airborne solid or liquid droplets with diameters of 2.5 micrometers (μm) or less.  PM~2.5~ levels tend to be highest around busy roads, industrial operations and neighbourhoods with residential wood burning. PM~2.5~ can easily penetrate indoors because of their small size. Common sources of PM~2.5~ that contribute to episodes of poor air quality  vary seasonally but can include wood smoke (from wood stoves and/or open burning) as well as emissions from industry and transportation sources such as automobiles, trucks and rail traffic.\n\n", "The provincial air quality objective for PM~2.5~ is 25 micrograms per cubic metre (μg/m^3^) averaged over 24 hours. 24-hour average PM~2.5~ concentrations are summarized below for ", params$nearestMonitor, " and nearest monitored communities at ", string_local_hour, " local time today:"
)
```

```{r burn_ban, results='asis', echo=FALSE}

# Define signature and title info depending on who it is
  if (params$burnRestrictions == 1) {
    name_line <- "Benjamin Weinstein"
    title_line1 <- "For Director, Environmental Management Act"
    title_line2 <- "Environmental Monitoring and Analysis Branch"
  } else if (params$burnRestrictions == 2) {
    name_line <- "TBD"
    title_line1 <- "Director, Environmental Management Act"
    title_line2 <- "Environmental Monitoring and Analysis Branch"
  }

# to do: improve formatting - single line spacing for sdm information, \n and <br> don't seem to work; rather than using paste could this text be added to a conditional code block directly in the document?
burn_ban <- paste0(
  "NOTICE of DIRECTOR under s.30 Open Burning Smoke Control Regulation:\n\n",
  "> As pollution is occurring or is likely to occur from open burning, pursuant to Sections 30(1) and 30(2) of the Open Burning Smoke Control Regulation, the Director has prohibited open burning within ",
  params$burnRestrictionArea,
  " until ",
  format(as.Date(params$burnRestrictionEndDate), "%B %d, %Y"),
  " ",
  params$burnRestrictionEndTime,
  " local time. No vegetative debris may be ignited or added to ignited piles. Contravention of these provisions may be subject to a fine under the Regulation.\n\n",
  "&emsp;&emsp;&nbsp;Date issued: ", format(Sys.Date(), "%B %d, %Y"), markup_separator,
  "&emsp;&emsp;&nbsp;", name_line, markup_separator,
  "&emsp;&emsp;&nbsp;", title_line1, markup_separator,
  "&emsp;&emsp;&nbsp;", title_line2
  )

```

```{r create-logo-list}
# Identify which set of logos to include in document header (note: BC Gov and FNHA logos are included in all cases so are not used to define logo file name)
# This list contains the full name (used for Alt Text and identification) on the left as the keys of the list and the initials (used for the logo file name) on the right as values of the list
logos <- construct_logos(params$outputFormat, string_health_authority)
```

<!-- Logo header, the layout-col should be set based on number of logos including FHNA and BCGov -->
<!-- the trailing slash means the text in square brackets is alt text -->

[`r paste("::: {layout-ncol=", logos$n_logos," layout-valign=\"bottom\"}")`]{.content-visible when-format="markdown"}

```{r print-logos}
#| results: asis

# `cat` is used to avoid extra processing
# `sep` argument adds the line break and then a new line as required
if (params$outputFormat == "markdown") cat(logos$logo_image_line, sep="\n\n")
```

[`r paste(":::")`]{.content-visible when-format="markdown"}

`r first_paragraph`

```{r, results='asis', echo=FALSE}
cat(
    "Open burning restrictions are now in effect within ", params$burnRestrictionArea,
    ". No new fires may be initiated, and no additional material may be added to existing fires. For more information on burning restrictions, refer to the Mandatory Emission Reduction Actions section below.",
    sep = ""
  )
```

Exposure to `r string_pollutant_expansion` is particularly a concern for infants, older adults, individuals with chronic
conditions (such as asthma, COPD, heart disease, and diabetes) or respiratory infections, and those who are pregnant. Anyone
experiencing symptoms such as continuing eye or throat irritation, chest discomfort,
shortness of breath, cough or wheezing, should follow the advice of their health care provider. Staying indoors helps to
reduce exposure.

`r params$customMessage`

The next update will be on `r format(as.Date(params$nextUpdate), "%B %d, %Y")` and posted to the
province’s [Air Quality Warnings webpage](https://www.gov.bc.ca/airquality).

Visit the
provincial [air quality data webpage](https://www2.gov.bc.ca/gov/content/environment/air-land-water/air/air-quality) for
real-time observations.

## Actions you can take

Exposure to poor air quality can affect your health. Reduce the duration and intensity of outdoor activities.
Spend time indoors in a space with filtered air.

Some people are more likely to be negatively impacted by outdoor air pollution. This includes people with lung
and heart conditions, pregnant individuals, infants and young children, people aged 65 and older, and anyone whose
health is affected by a chronic condition or an acute illness.

::: {.content-visible when-format="markdown"}
{{< accordion_controls >}}
{{< accordion_start title="Follow your common sense" initiallyOpen="true" >}}
:::

::: {.content-visible when-format="pdf"}
### Follow your common sense
:::

- Stop or reduce your activity level if breathing becomes uncomfortable or you feel unwell.

```{r, results='asis', echo=FALSE}

if (current_month %in% 6:8) {
  cat("- During warm weather, stay cool and drink plenty of fluids.", markup_separator)
}
```

- Always carry any rescue medications with you.
- Make sure that children and others who cannot care for themselves follow the same advice.

::: {.content-visible when-format="markdown"}
{{< accordion_end >}}
{{< accordion_start title="Monitor your symptoms" initiallyOpen="true" >}}
:::

::: {.content-visible when-format="pdf"}
### Monitor your symptoms
:::

- Different people have different responses to elevated levels of air contaminants.
- Mild irritation and discomfort such as eye, nose and throat irritation, headaches or a mild cough are common, and
  usually disappear when the air contaminants return to typical levels.
- More serious but less common symptoms include wheezing, chest pains or severe cough.
- People with asthma or other chronic illness should follow any personal care plans designed with their family
  physicians.
- If you are unsure whether you need medical care, call HealthLink BC at 8-1-1.
- If you are experiencing difficulty in breathing, chest pain or discomfort, or a severe cough, contact your physician,
  walk-in clinic, or emergency department. **If you are having a medical emergency, call 9-1-1.**

::: {.content-visible when-format="markdown"}
{{< accordion_end >}}
{{< accordion_start title="Tips to reduce your exposure to air pollution" initiallyOpen="true" >}}
:::

::: {.content-visible when-format="pdf"}
### Tips to reduce your exposure to air pollution
:::

- Air contaminant levels may be elevated indoors, so stay aware of your symptoms even when you are indoors.

```{r, results='asis', echo=FALSE}

if (current_month %in% 6:8) {
  cat(
    "- When indoors, keep windows and doors closed.  If it is hot, prioritize keeping cool.\n",
    "- When there is an extreme heat event ocurring with poor air quality , prioritize keeping cool.\n",
    "- If your home is too hot, spend time in places with air conditioning such as community centres, libraries, or malls.\n"
    )
}

if (current_month %in% c(1:5, 9:12)) {
  cat("- When indoors, keep windows and doors closed.\n")
}

cat(
    "- Protect your indoor air from outdoor air pollution. Actions can include using a clean, good quality air filter in your ventilation system and/or a certified portable air cleaner that can filter fine particles. Do-it-yourself air cleaners may also be used if other options are unavailable. For more details, see the BC Centre of Control [fact sheet](https://www.bccdc.ca/resource-gallery/Documents/Guidelines%20and%20Forms/Guidelines%20and%20Manuals/Health-Environment/BCCDC_WildFire_FactSheet_BoxFanAirFilters.pdf).\n",
    "- If you must spend time outdoors, a well-constructed, well-fitting and properly worn respirator type mask (such as a NIOSH-certified N95 or equivalent respirator) can reduce your exposure to particulate matter. Even though exposure may be reduced, there can still be risks to health.\n"
  )
```

::: {.content-visible when-format="markdown"}
{{< accordion_end >}}
:::

## Emission reduction actions

::: {.content-visible when-format="markdown"}
{{< accordion_start title="Mandatory actions" initiallyOpen="true" >}}
`r if (params$burnRestrictions > 0) paste("-", burn_ban)`
{{< accordion_end >}}
:::

::: {.content-visible when-format="pdf"}
### Mandatory actions
`r if (params$burnRestrictions > 0) paste("-", burn_ban)`
:::

::: {.content-visible when-format="markdown"}
{{< accordion_start title="Voluntary actions" initiallyOpen="true" >}}
:::

::: {.content-visible when-format="pdf"}
### Voluntary actions
:::

`r voluntary_reduction`

::: {.content-visible when-format="markdown"}
{{< accordion_end >}}
:::

## More information

::: {.content-visible when-format="markdown"}
{{< accordion_start title="Current conditions" initiallyOpen="true" >}}
:::

::: {.content-visible when-format="pdf"}
### Current conditions
:::

```{r pollutants-table, results='asis', echo=FALSE}
# Single pollutant: PM25
  cat(additional_info_PM25, "\n\n")
  cat('[::: {.table-responsive}]{.content-visible when-format="markdown"}\n')
  print(knitr::kable(data_pollutants_table, row.names = FALSE, escape = FALSE))
  cat('[\n:::]{.content-visible when-format="markdown"}\n')
```

::: {.content-visible when-format="markdown"}
{{< accordion_end >}}
:::

::: {.content-visible when-format="pdf"}
### For additional information about air quality in British Columbia
:::

::: {.content-visible when-format="markdown"}
**For additional information about air quality in British Columbia:**

::: bcds-card-wrapper
{{< card_start title="Air Quality Health Index" variant="info" logo="/assets/icon_air.svg" >}} Provincial summary of
latest index and forecast.

[What's the air like today?](https://www.env.gov.bc.ca/epd/bcairquality/data/aqhi-table.html) {{< card_end >}} {{<
card_start title="Provincal Air Quality Map" variant="info" logo="/assets/icon_air.svg" >}} Provincial air quality data
refreshed every hour.

[Latest air quality data](https://www.env.gov.bc.ca/epd/bcairquality/readings/find-stations-map.html) {{<
card_end >}} {{< card_start title="Metro Vancouver Air Quality" variant="success" logo="/assets/icon_chart.svg" >}}
Warnings and a map of air quality and weather data.

[Air quality data and warnings](https://metrovancouver.org/services/air-quality-climate-action/air-quality-data-and-advisories) {{<
card_end >}}
  ```{r, results='asis', echo=FALSE, eval= run_pg}

  cat('{{< card_start title="Local Resources" variant="info" logo="/assets/icon_air.svg" >}}\n\n [Prince George Air Improvement Roundtable](https://www.pgairquality.com/) {{< card_end >}}')
  ```
:::
:::

::: {.content-visible when-format="pdf"}
- View the latest [Air Quality Health Index (AQHI) and forecast](https://www.env.gov.bc.ca/epd/bcairquality/data/aqhi-table.html)
- Visit the provincial [air quality map](https://www.env.gov.bc.ca/epd/bcairquality/readings/find-stations-map.html) to view real-time air quality measurements
- Visit the provincial [air quality data webpage](https://www2.gov.bc.ca/gov/content/environment/air-land-water/air/air-quality) for more information about air quality in B.C.
- Visit Metro Vancouver's [air quality data and warnings page](https://metrovancouver.org/services/air-quality-climate-action/air-quality-data-and-advisories)
- Subscribe to receive air quality warnings on the [Air Quality Subscriptions Service](https://www2.gov.bc.ca/gov/content/environment/air-land-water/air/air-quality/air-advisories/air-quality-subscription-service) page
```{r, results='asis', echo=FALSE, eval= run_pg}
  cat("- Visit the [Prince George Air Improvement Roundtable's page](https://www.pgairquality.com/)")
```
:::

## Contact

::: {.content-visible when-format="pdf"}
### Media questions regarding this Pollution Prevention Notice
:::

::: {.content-visible when-format="markdown"}
**Media questions regarding this Pollution Prevention Notice:**
{{< card_start width="wide" >}}
:::

`r get_meteorologist_contact(params$aqMet, markup_separator)`

::: {.content-visible when-format="markdown"}
{{< card_end >}}
:::

::: {.content-visible when-format="pdf"}
### Media questions regarding health implications when air quality is degraded
:::

::: {.content-visible when-format="markdown"}
**Media questions regarding health implications when air quality is degraded:**
{{< card_start width="wide" >}}
:::

`r paste(data_health_authority_contact$html_string, collapse = "\n\n")`

::: {.content-visible when-format="markdown"}
{{< card_end >}}
:::
