---
date: "`r Sys.Date()`"
params:
  aqMet: "Sakshi Jain"
  lastWarning: "2025-02-10"
  customMessage: "`Custom message.`"
  location: "Multiple locations in B.C."
  outputFormat: "markdown"
  healthAuth:
    - First Nations Health Authority
    - Northern Health
format:
  markdown: default
  pdf:
    mainfont: "Roboto"
    date-format: long
    include-in-header:
      - src/header.tex
    fontsize: "12pt"
    urlcolor: bcblue
---

```{=html}
<!--
Copyright 2025 Province of British Columbia

This work is licensed under the Creative Commons Attribution 4.0 International License.
To view a copy of this license, visit http://creativecommons.org/licenses/by/4.0/.
-->
```

```{r}
#| label: setup
#| echo: FALSE
#| include: FALSE
#| warning: FALSE
#| message: FALSE

# Set options globally
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

# Load packages
library(dplyr)

# Import files in the quarto environment
source(here::here("load_metadata.r"))
source(here::here(file.path("src", "construct_logos.r")))
source(here::here(file.path("src", "construct_shared_dependencies.r")))

```

```{r}
#| label: lookup
#| results: asis
#| strip.white: false

if (params$outputFormat == "markdown") markup_separator <- "<br />" else markup_separator <- "  \n"

#AQ met information
data_health_authority_contact <- reference_health_authority_contact |>
  filter(authority %in% params$healthAuth) |>
  mutate(authority = factor(authority))|>
  select(authority, contact) |>
  group_by(authority) |>
  summarise(html_string = paste(contact, collapse = markup_separator))

```

```{r}
#| label: create-logo-list
logos <- construct_logos(params$outputFormat, params$healthAuth)
```

<!-- Logo header, the layout-col should be set based on number of logos including FHNA and BCGov -->
<!-- the trailing slash means the text in square brackets is alt text -->

[`r paste("::: {layout-ncol=", logos$n_logos," layout-valign=\"bottom\"}")`]{.content-visible when-format="markdown"}

```{r}
#| label: print-logos
#| results: asis

# `cat` is used to avoid extra processing
# `sep` argument adds the line break and then a new line as required
if (params$outputFormat == "markdown") cat(logos$logo_image_line, sep="\n\n")
```

[`r paste(":::")`]{.content-visible when-format="markdown"}

The Air Quality Warning for wildfire smoke that was last updated on `r format(as.Date(params$lastWarning), '%B %d, %Y')` has ended.

`r params$customMessage`

Real-time air quality information is available on the [BC Air Quality website](https://gov.bc.ca/airqualityhealthindex).

# Contact information

**Media questions regarding air quality conditions related to this Air Quality Warning:**

[{{< card_start width="wide" >}}]{.content-visible when-format="markdown"}
`r get_meteorologist_contact(params$aqMet, markup_separator)`
[{{< card_end >}}]{.content-visible when-format="markdown"}

**Media questions regarding health implications of wildfire smoke:**

[{{< card_start width="wide" >}}]{.content-visible when-format="markdown"}`r paste(data_health_authority_contact$html_string, collapse = "\n\n")`[{{< card_end >}}]{.content-visible when-format="markdown"}
