---
title: "Air quality warning for wildfire smoke ended" 
type: "wildfire_smoke" 
date: "`r Sys.Date()`" 
params: 
  aqMet: "Sakshi Jain" 
  lastWarning: "2025-02-10" 
  customMessage: "`Custom message.`" 
  location: "Multiple locations in B.C."
  outputFormat: "markdown"
  healthAuth: 
    - First Nations Health Authority 
    - Northern Health 
ice: "End" 
author: "`r params$aqMet`" 
customMessage: "`r params$customMessage`" 
lastWarning: "`r params$lastWarning`" 
healthAuth: "`r paste(params$healthAuth, collapse=', ')`" 
location: "`r params$location`"
outputFormat: "`r params$outputFormat`" 
format:
  markdown: default
  pdf:
    mainfont: "Roboto"
    date-format: long
    include-in-header: 
      - src/header.tex
    fontsize: "12pt"
    urlcolor: bcblue
---

```{=html}
<!--
Copyright 2025 Province of British Columbia

This work is licensed under the Creative Commons Attribution 4.0 International License.
To view a copy of this license, visit http://creativecommons.org/licenses/by/4.0/.
-->
```

```{r}
#| label: setup
#| echo: FALSE
#| include: FALSE
#| warning: FALSE
#| message: FALSE

# Set options globally
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

# Load packages
library(dplyr)

# Import files in the quarto environment
source(here::here("load_metadata.r"))

```

```{r}
#| label: lookup
#| results: asis
#| strip.white: false

if (params$outputFormat == "markdown") sep_type <- "<br />" else sep_type <- "  \n"

#AQ met information

ENVcontact <- aq_mets |>
  filter(fullname == params$aqMet) |>
  mutate(contact = paste(fullname_typeset, title, ministry, phone, sep = sep_type)) |> 
  pull(contact)

HAcontact <- health_contact |>
  filter(authority %in% params$healthAuth) |>
  mutate(authority = factor(authority))|> 
  select(authority, contact) |>
  group_by(authority) |> 
  summarise(html_string = paste(contact, collapse = sep_type))

```

```{r}
#| label: create-logo-list

# Identify which set of logos to include in document header (note: BC Gov and FNHA logos are included in all cases so are not used to define logo file name)
# This list contains the full name (used for Alt Text and identification) on the left as the keys of the list and the initials (used for the logo file name) on the right as values of the list

logos_list <- list(
  "Government of British Columbia" = "BCID_V_RGB_pos",
  "First Nations Health Authority" = "FNHA",
  "Interior Health Authority" = "IH", 
  "Fraser Health Authority" = "FH", 
  "Vancouver Coastal Health Authority" = "VCH", 
  "Vancouver Island Health Authority" = "VIH", 
  "Northern Health Authority" = "NH")

# Logos selected by user and ordered as per logos_list
logos_names_selected <- c(
  "Government of British Columbia", # always select BC
  "First Nations Health Authority", # always select FNHA 
  params$healthAuth) # Note: if FNHA is selected in app, it will duplicate

## This line will remove the duplicated FNHA value if it appears twice
## Side effect: If the app user did NOT select FNHA, this will force FNHA to appear
logos_names_selected <- unique(logos_names_selected)
logos_names_selected <- logos_names_selected[order(match(logos_names_selected, names(logos_list)))]  # match order to logos_list
logos_selected <- logos_list[logos_names_selected] # subset of logo_list

# Count number of logos to display
n_logos <- length(logos_selected)

logo_image_line <- c()
logos_combined <- c()

# Set logo path based on output format
if (params$outputFormat == "markdown") {
  logo_path <- " logo](//assets/logo_"

  # Build a vector of quarto lines to insert image for each logo
  # It is more efficient to use `sapply` but this might be more readable

  # Add each logo's insert line to the vector
  for (logo_name in names(logos_selected)) {
    logo_image_line <- c(logo_image_line,
                         paste0("![", logo_name, logo_path, logos_selected[[logo_name]], ".png)\\"))
  }}


if (params$outputFormat == "pdf") {
  
  logo_path <- "https://github.com/bcgov/aqwarnings/blob/main/frontend/assets/logo_"
  logo_urls <- c() #start with empty vector
  
  # create vector of urls
  for (logo_name in names(logos_selected)) {
    logo_urls <- c(
      logo_urls, 
      paste0(logo_path,
             logos_selected[[logo_name]],
             ".png?raw=true"
             )
    )
  }

# format logos
logos <- magick::image_read(logo_urls) |> 
  magick::image_trim() |> # remove all white space
  magick::image_border(color = "none", geometry = "100x25") |> # add uniform white space to top and right side of each image
  magick::image_scale("x200") # scale

# Combine logos into single image and scale
logos_combined <- magick::image_append(logos) |> 
  magick::image_write(path = "logo.png")
}

```

<!-- Logo header, the layout-col should be set based on number of logos including FHNA and BCGov -->
<!-- the trailing slash means the text in square brackets is alt text -->

[`r paste("::: {layout-ncol=", n_logos," layout-valign=\"bottom\"}")`]{.content-visible when-format="markdown"}

```{r}
#| label: print-logos
#| results: asis

# `cat` is used to avoid extra processing
# `sep` argument adds the line break and then a new line as required
if (params$outputFormat == "markdown") cat(logo_image_line, sep="\n\n") 
```

[`r paste(":::")`]{.content-visible when-format="markdown"}

This Air Quality Warning for wildfire smoke that was last updated on `r format(as.Date(params$lastWarning), '%B %d, %Y')` has ended.

`r params$customMessage`

Real-time air quality information is available on the [BC Air Quality website](https://gov.bc.ca/airqualityhealthindex).

# Contact information

**Media questions regarding air quality conditions related to this Air Quality Warning:**

[{{< card_start width="wide" >}}]{.content-visible when-format="markdown"}`r ENVcontact`[{{< card_end >}}]{.content-visible when-format="markdown"}

**Media questions regarding health implications of wildfire smoke:**

[{{< card_start width="wide" >}}]{.content-visible when-format="markdown"}`r paste(HAcontact$html_string, collapse = "\n\n")`[{{< card_end >}}]{.content-visible when-format="markdown"}
