---
title: "Air Quality Warning - Wildfire Smoke"
type: "wildfire_smoke" 
date: "`r Sys.Date()`"
params:
  sel_aqMet: "Sakshi"
  smokeDuration: "`24-48 hours`"
  customMessage: "`Custom message.`"
  ice: "Issue"
  nextUpdate: "2025-02-10"
  location: "Multiple locations in B.C."
  selRegionsIDs:
    - Watson Lake
    - Prince George
    - Whistler
    - South Okanagan
ice: "`r params$ice`"
author: "`r params$sel_aqMet`"
smokeDuration: "`r params$smokeDuration`"
customMessage: "`r params$customMessage`"
location: "`r params$location`"
---

```{=html}
<!--
Copyright 2025 Province of British Columbia

This work is licensed under the Creative Commons Attribution 4.0 International License.
To view a copy of this license, visit http://creativecommons.org/licenses/by/4.0/.
-->
```

```{r setup, echo=FALSE, include=FALSE, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(echo=FALSE,
                      message=FALSE,
                      warning=FALSE)

library(dplyr)
library(leaflet)

#Import files in the quarto environment
source(here::here("load_metadata.r"))

```

```{r lookup, results="asis", strip.white=FALSE}
#AQ met information

#this may not be needed? test later
ENVcontact <- aq_mets |>
  filter(fullname == params$sel_aqMet) |>
  mutate(contact = paste(fullname_typeset, title, ministry, phone, sep = "<br />")) |> 
  pull(contact)

# Check for first health authority
a <- match_eccc_health |> 
  filter(eccc %in% params$selRegionsIDs) |> 
  distinct(healthAuth1) |> 
  pull(healthAuth1)

# Check for second health authority
b <-  match_eccc_health |> 
  filter(eccc %in% params$selRegionsIDs) |> 
  distinct(healthAuth2) |> 
  pull(healthAuth2)

# find unique health authorities between a and b; drop empty cases (""")
unique_HA <- unique(c(a,b))
unique_HA <- unique_HA [! unique_HA %in% ""]

# filter health authorities, always include FNHA
HAcontact <- health_contact |> 
   filter(authority %in% c("First Nations Health Authority", unique_HA)) |> 
   mutate(authority = factor(authority))

# retrieve descriptions (utility_files/eccc_descriptions.csv) for each affected region
# descriptions included at end of bulletin
bulletin_regions <- reg_description |> 
  filter(eccc %in% params$selRegionsIDs) |> 
  filter(!is.na(region_name) & region_name != "")|> 
  mutate(eccc = factor(eccc))

region <- paste(bulletin_regions$region_name, collapse = ", ")

description <- paste(bulletin_regions$description, collapse = "; ")

region_output <- bulletin_regions |> 
  distinct(region_name, description) |> 
  mutate(output_line = paste0("* ", region_name, ": ", description)) |> 
  pull(output_line) |> 
  paste(collapse = "\n")

```

```{r}
#| results: asis

initial_lat = 54.8
initial_long = -124.253144

currentDate <- Sys.Date()
issueBasename <- "wildfire_smoke_issue"

#adding a dynamic map to the html file using area boundaries from ECCC. 
m <- leaflet() |> 
  addTiles() |> 
  addPolygons(
        data = eccc_map_env,
        fillOpacity = 0.025,
        opacity = 0.75,
        color = "black",
        fillColor = "#FFF716",
        stroke = TRUE,
        weight = 1,
        smoothFactor = 0.2,
        layerId = ~NAME,
        group = "regions",
        label = ~NAME,
        labelOptions = labelOptions(textsize = "15px"),
        ) |> 
  addPolygons(data = eccc_map_env[which(eccc_map_env$NAME %in% params$selRegionsIDs), ],
                  fillOpacity = 0.65,
                  opacity = 0.65,
                  color = "black",
                  weight = 1.75,
                  fillColor = "#FFF716",
                  label = ~NAME,
                  labelOptions = labelOptions(textsize = "15px"))

map_html <- sprintf(file.path("%s_%s_map.html"), currentDate, issueBasename)
m <- mapview::mapshot(m, url = map_html, cliprect = c(140, 147, 610, 480))
```

```{r}
# Add in extra information if Interior Health Authority (IH) is present in contacts
HA_text <- if (any(HAcontact$authority == "Interior Health Authority")) {
  "{{< card_start title=\"Interior Health Authority\" logo=\"/assets/logo_IH_horizontal.png\" >}}
Steps to prepare and manage wildfire-related issues.

[Wildfire public information](https://www.interiorhealth.ca/health-and-wellness/natural-disasters-and-emergencies/wildfires)
{{< card_end >}}"
} else {
  NULL
}

```

```{r}
# Identify which set of logos to include in document header (note: BC Gov and FNHA logos are included in all cases so are not used to define logo file name)
# This list contains the full name (used for Alt Text and identification) on the left as the keys of the list and the initials (used for the logo file name) on the right as values of the list
logos_list <- list(
  "Government of British Columbia" = "BCID_V_RGB_pos",
  "First Nations Health Authority" = "FNHA",
  "Interior Health Authority" = "IH", 
  "Fraser Health Authority" = "FH", 
  "Vancouver Coastal Health Authority" = "VCH", 
  "Vancouver Island Health Authority" = "VIH", 
  "Northern Health Authority" = "NH")

# Logos selected by user and ordered as per logos_list
logos_names_selected <- c("Government of British Columbia", "First Nations Health Authority",  unique_HA[order(match(unique_HA, names(logos_list)))]) # always select BC and FNHA and sort unique_HA to match logos_list
logos_selected <- logos_list[logos_names_selected] # subset of logo_list

# Count number of logos to display
n_logos <- length(logos_selected)

# Build a vector of quarto lines to insert image for each logo
# It is more efficient to use `sapply` but this might be more readable
logo_image_line <- c() # start with empty vector
# Add each logo's insert line to the vector
for(logo_name in names(logos_selected)) {
   logo_image_line <- c(logo_image_line, paste0("![", logo_name, " logo](//assets/logo_", logos_selected[[logo_name]], ".png)\\"))
}
```

<!-- Logo header, the layout-col should be set based on number of logos including FHNA and BCGov-->

<!-- the trailing slash means the text in square brackets is alt text -->

`r paste("::: {layout-ncol=", n_logos," layout-valign=\"bottom\"}")`

```{r}
#| results: asis
# `cat` is used to avoid extra processing
# `sep` argument adds the line break and then a new line as required
cat(logo_image_line, sep="\n\n")
```

`r paste(":::")`

The Regions of BC highlighted on the map are being impacted or are likely to be impacted by wildfire smoke over the next `r params$smokeDuration`.

`r params$customMessage`

The next update will be available on `r format(as.Date(params$nextUpdate), '%B %d, %Y')`.

The Air Quality Warning for wildfire smoke can be accessed online at [https://bcgov.github.io/aqwarnings/](https://bcgov.github.io/aqwarnings/).

![A description of Region(s) included in this Air Quality Warning is provided at the end of this page. This Air Quality Warning excludes the area managed by Metro Vancouver. Refer to the More Information section for a link to air quality notifications issued by Metro Vancouver.](`r map_html`){fig-alt="Air Quality Warning - Wildfire Smoke Regions for `r format(Sys.Date(), '%B %d, %Y')`: `r region`." width="100%" height="600px"}

## Be informed

During a wildfire, smoke conditions can change quickly over short distances and can vary considerably hour-by-hour.

Wildfire smoke is a natural part of our environment but it is important to be mindful that exposure to smoke may affect your health.

People with pre-existing health conditions, respiratory infections such as COVID-19, older adults, pregnant women and infants, children, and sensitive individuals are more likely to experience health effects from smoke exposure.

## During smoky conditions

{{< accordion_controls >}}

{{< accordion_start title="Follow your common sense" initiallyOpen="true" >}}

-   Stop or reduce your activity level if breathing becomes uncomfortable or you feel unwell.

-   Stay cool and drink plenty of fluids.

-   If you have asthma or other chronic illness, carry any rescue (fast-acting) medications with you at all times and activate your personal care plan that has been designed with your family physician.

-   Make sure that children and others who cannot care for themselves follow the same advice.

{{< accordion_end >}}

{{< accordion_start title="Monitor your symptoms" initiallyOpen="true" >}}

-   People respond differently to smoke. Mild irritation and discomfort are common, and usually disappear when the smoke clears.

-   If you are unsure whether you need medical care, call HealthLink BC at 8-1-1.

-   If you are experiencing difficulty in breathing, chest pain or discomfort, or a severe cough, contact your health care provider, walk-in clinic, or emergency department. If you are having a medical emergency, call 9-1-1.

{{< accordion_end >}}

{{< accordion_start title="Tips to reduce your smoke exposure" initiallyOpen="true" >}}

-   Smoke levels may be lower indoors but will still be elevated, so stay aware of your symptoms even when you are indoors.

-   Running a commercially available HEPA (high efficiency particulate air) filter can improve indoor air quality in the room where the device is located.

-   If you have a forced air heating/cooling system in your home, it may help to change the filter and set the fan to run continuously.

-   Reduce indoor air pollution sources such as smoking, burning incense, and frying foods.

-   Consider going to a library, community center, or shopping mall with cooler filtered air to get some relief from the smoke.

-   If travelling in a car with air conditioning, keep the windows up and the ventilation set to recirculate.

-   If you are very sensitive to smoke, consider moving to another location with cleaner air, but be aware that conditions can change rapidly.

-   Maintaining good overall health is a good way to prevent health effects resulting from short-term exposure to air pollution.

{{< accordion_end >}}

## More information

**For additional general information about wildfire smoke and air quality:**

<div class="bcds-card-wrapper"> 
{{< card_start title="Air Quality Health Index" variant="info" logo="/assets/icon_air.svg" >}} Provincial summary of latest index and forecast.

[What's the air like today?](https://www.env.gov.bc.ca/epd/bcairquality/data/aqhi-table.html) {{< card_end >}} {{< card_start title="Air Quality Map" variant="info" logo="/assets/icon_air.svg" >}} Provincial air quality data refreshed every hour.

[Latest air quality data](https://www.env.gov.bc.ca/epd/bcairquality/readings/find-stations-map.html) {{< card_end >}} {{< card_start title="Environment and Climate Change Canada FireWork" variant="danger" logo="/assets/icon_smoke.svg" >}} Air quality maps of next 72 hours.

[Air quality forecast maps](https://weather.gc.ca/firework/index_e.html) {{< card_end >}} {{< card_start title="FireSmoke Canada" variant="danger" logo="/assets/icon_smoke.svg" >}} BlueSky Canada smoke forecasting system.

[Smoke forecasts](https://firesmoke.ca/forecasts/current/) {{< card_end >}} {{< card_start title="Metro Vancouver Air Quality" variant="success" logo="/assets/icon_chart.svg" >}} Advisories and map of air quality and weather data.

[Air quality data and advisories](https://metrovancouver.org/services/air-quality-climate-action/air-quality-data-and-advisories) {{< card_end >}}
</div>

**For additional general information about wildfire smoke and your health:**

<div class="bcds-card-wrapper"> 
{{< card_start title="BC Centre for Disease Control" logo="/assets/logo_BCCDC.png" >}} Fact sheets on wildfire smoke and its health impacts.

[Wildfire smoke](http://www.bccdc.ca/health-info/prevention-public-health/wildfire-smoke) {{< card_end >}}`r HA_text`{{< card_start title="First Nations Health Authority" logo="/assets/logo_FNHA_horizontal.png" >}} Information and resources to help you prepare.

[Wildfire smoke health information](https://www.fnha.ca/about/news-and-events/news/wildfire-smoke-health-information) {{< card_end >}} {{< card_start title="HealthLink BC" logo="/assets/logo_BCID_HealthLinkBC_RGB_pos.png" >}} Learn and prepare for wildfires. Phone 8-1-1 for 24/7 health advice.

[Wildfires and your health](https://www.healthlinkbc.ca/health-library/health-features/wildfires-and-your-health) {{< card_end >}} {{< card_start title="Worksafe BC" logo="/assets/logo_worksafebc.jpg" >}} Info on working outdoors during smoky conditions.

[Wildfire FAQ](https://www.worksafebc.com/en/resources/health-safety/information-sheets/wildfire-smoke-frequently-asked-questions-faq) {{< card_end >}}
</div>

## Contact

**Media and public inquiries regarding air quality and the Air Quality Warning for wildfire smoke:**

{{< card_start width="wide" >}}
`r ENVcontact`
{{< card_end >}}

**Media questions regarding health implications of wildfires:**

{{< card_start  width="wide" >}}
`r paste(HAcontact$contact, collapse = "<br />")`
{{< card_end >}}


{{< accordion_start title="Regions included under the Air Quality Warning - Wildfire Smoke" headerClass="h2" >}}

`r paste(region_output)`

{{< accordion_end >}}
