---
date: "`r Sys.Date()`"
params:
  aqMet: "Sakshi Jain"
  smokeDuration: "`24-48 hours`"
  customMessage: "`Custom message.`"
  nextUpdate: "2025-02-10"
  location: "Multiple locations in B.C."
  outputFormat: "markdown"
  warningLevel:
    value:
      all:
        - Watson Lake
        - Prince George
        - Whistler
        - South Okanagan
      yellow:
        - Watson Lake
        - Prince George
      orange:
        - Whistler
      red:
        - South Okanagan
  twoLevels: FALSE
format:
  markdown: default
  pdf:
    mainfont: "Roboto"
    date-format: long
    include-in-header:
      - src/header.tex
    fontsize: "12pt"
    urlcolor: bcblue
---

```{=html}
<!--
Copyright 2025 Province of British Columbia

This work is licensed under the Creative Commons Attribution 4.0 International License.
To view a copy of this license, visit http://creativecommons.org/licenses/by/4.0/.
-->
```

```{r}
#| label: setup
#| echo: FALSE
#| include: FALSE
#| warning: FALSE
#| message: FALSE

# Set options globally
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

# Load packages
library(dplyr)

# Import files in the quarto environment
source(here::here("load_metadata.r"))
source(here::here(file.path("src", "construct_banners.r")))
source(here::here(file.path("src", "construct_logos.r")))
source(here::here(file.path("src", "construct_shared_dependencies.r")))

```

```{r}
#| label: lookup
#| results: asis
#| strip.white: false

if (params$outputFormat == "markdown") markup_separator <- "<br />" else markup_separator <- "  \n"

# Check for first health authority
a <- reference_match_eccc_health |>
  filter(eccc %in% params$warningLevel$all) |>
  distinct(healthAuth1) |>
  pull(healthAuth1)

# Check for second health authority
b <- reference_match_eccc_health |>
  filter(eccc %in% params$warningLevel$all) |>
  distinct(healthAuth2) |>
  pull(healthAuth2)

# Find unique health authorities between a and b; drop empty cases (" ")
unique_HA <- unique(c(a,b))
unique_HA <- unique_HA [! unique_HA %in% ""]

# Filter health authorities, always include FNHA
data_health_authority_contact <- reference_health_authority_contact |>
  filter(authority %in% c("First Nations Health Authority", unique_HA)) |>
  mutate(authority = factor(authority))|>
  select(authority, contact) |>
  group_by(authority) |>
  summarise(html_string = paste(contact, collapse = markup_separator))

# Retrieve descriptions (utility_files/eccc_descriptions.csv) for each affected region
# descriptions included at end of bulletin
local_bulletin_regions <- reference_reg_description |>
  filter(eccc %in% params$warningLevel$all) |>
  filter(!is.na(region_name) & region_name != "")|>
  mutate(eccc = factor(eccc))

string_region_names <- paste(local_bulletin_regions$region_name, collapse = ", ")

string_region_output <- local_bulletin_regions |>
  distinct(region_name, description) |>
  mutate(output_line = paste0("* **", region_name, ":** ", description)) |>
  pull(output_line) |>
  paste(collapse = "\n")

flag_run_pg <- "Prince George" %in% local_bulletin_regions$region
```

```{r}
#| label: add-map
#| results: asis

initial_lat = 54.8
initial_long = -124.253144

currentDate <- Sys.Date()
issueBasename <- "wildfire_smoke_issue"

# Add dynamic map to quarto environment using area boundaries from ECCC.
m <- leaflet::leaflet() |>
  leaflet::addTiles() |>
  leaflet::addPolygons(
    data = reference_eccc_map_env,
    fillOpacity = 0.025,
    opacity = 0.75,
    color = "black",
    fillColor = "grey",
    stroke = TRUE,
    weight = 1,
    smoothFactor = 0.2,
    layerId = ~ NAME,
    group = "regions",
    label = ~ NAME,
    labelOptions = leaflet::labelOptions(textsize = "15px"),
  )

  if (length(params$warningLevel$yellow) > 0) {
    m <- leaflet::addPolygons(m,
      data = reference_eccc_map_env[which(reference_eccc_map_env$NAME %in% params$warningLevel$yellow), ],
      fillOpacity = 0.65,
      opacity = 0.65,
      color = "black",
      weight = 1.75,
      fillColor = "#FFFF00",
      label = ~ NAME,
      labelOptions = leaflet::labelOptions(textsize = "15px")
    )
  }

  if (length(params$warningLevel$orange) > 0) {
      m <- leaflet::addPolygons(m,
        data = reference_eccc_map_env[which(reference_eccc_map_env$NAME %in% params$warningLevel$orange), ],
        fillOpacity = 0.65,
        opacity = 0.65,
        color = "black",
        weight = 1.75,
        fillColor = "#FF9500",
        label = ~ NAME,
        labelOptions = leaflet::labelOptions(textsize = "15px")
      )
  }

  if (length(params$warningLevel$red) > 0) {
    m <- leaflet::addPolygons(m,
      data = reference_eccc_map_env[which(reference_eccc_map_env$NAME %in% params$warningLevel$red), ],
      fillOpacity = 0.65,
      opacity = 0.65,
      color = "black",
      weight = 1.75,
      fillColor = "#D10000",
      label = ~ NAME,
      labelOptions = leaflet::labelOptions(textsize = "15px")
    )
  }

map_html <- sprintf("%s_%s_map.html", currentDate, issueBasename)
map_png <- sprintf("%s_%s_map.png", currentDate, issueBasename)
m <- mapview::mapshot(m, url = map_html, cliprect = c(140, 147, 610, 480))

map <- if (params$outputFormat == "markdown") map_html else map_png
```

```{r}
#| label: add-ha-info
# Add in extra information if Interior Health Authority (IH) is present in contacts

if(params$outputFormat == "markdown") {
  HA_text <- if (any(data_health_authority_contact$authority == "Interior Health Authority")) {
  "{{< card_start title=\"Interior Health Authority\" logo=\"/assets/logo_IH_horizontal.png\" >}}
Steps to prepare and manage wildfire-related issues.

[Wildfire public information](https://www.interiorhealth.ca/health-and-wellness/natural-disasters-and-emergencies/wildfires)
{{< card_end >}}"
    }}

if (params$outputFormat == "pdf") {
      HA_text <- if (any(data_health_authority_contact$authority == "Interior Health Authority")) {
      "* Interior Health Authority - [Steps to prepare for and manage wildfire-related issues](https://www.interiorhealth.ca/health-and-wellness/natural-disasters-and-emergencies/wildfires)"
}}

```
```{r}
#| label: add-banners

banners <- ""

if (length(params$warningLevel$yellow) > 0) {
  if (params$outputFormat == "markdown") {
    banners <- paste(banners, "{{< banner_alert_start title='Yellow Warning - Air Quality' variant='yellow'>}}")
    banners <- paste(banners, "{{< banner_alert_end >}}")
  } else if (params$outputFormat == "pdf") {
    banners <- paste(banners, "\\YellowBanner")
  }
}

if (length(params$warningLevel$orange) > 0) {
  if (params$outputFormat == "markdown") {
    banners <- paste(banners, "{{< banner_alert_start title='Orange Warning - Air Quality' variant='orange'>}}")
    banners <- paste(banners, "{{< banner_alert_end >}}")
  } else if (params$outputFormat == "pdf") {
    banners <- paste(banners, "\\OrangeBanner")
  }
}

 if (length(params$warningLevel$red) > 0) {
  if (params$outputFormat == "markdown") {
    banners <- paste(banners, "{{< banner_alert_start title='Red Warning - Air Quality' variant='red'>}}")
    banners <- paste(banners, "{{< banner_alert_end >}}")
  } else if (params$outputFormat == "pdf") {
    banners <- paste(banners, "\\RedBanner")
  }
}
```
```{r}
#| label: add-warning-level-text
#| include: FALSE

# Create dynamic text based on warning level

string_all_levels <- "Some people are more likely to be negatively impacted by smoke. This includes people with lung and heart conditions, pregnant individuals, infants and young children, people aged 65 and older, and anyone whose health is affected by a chronic condition or acute illness."

string_yellow_text <- "Exposure to smoke can affect your health. Postpone or reduce strenuous outdoor activities until the warning has ended. Spend time indoors in a space with cleaner air."

string_orange_text <- "Exposure to high smoke levels can pose threat to your health regardless of age of health status. Avoid strenuous outdoor activities until the warning has ended. Stay indoors in a space with cleaner air."

string_action_text <- if(length(params$warningLevel$yellow) > 0 & length(params$warningLevel$orange) == 0) {
  string_yellow_text
} else if(!params$twoLevels & length(params$warningLevel$orange) > 0 & length(params$warningLevel$yellow) == 0) {
  string_orange_text
}
```
```{r}
#| label: create-logo-list
logos <- construct_logos(params$outputFormat, unique_HA[order(match(unique_HA, names(reference_logos_list)))])

```

<!-- Logo header for markdown output, the layout-col should be set based on number of logos including FHNA and BCGov -->
<!-- the trailing slash means the text in square brackets is alt text -->
<!-- note: logos and title for PDF output handled in src/header.txt -->

[`r paste("::: {layout-ncol=", logos$n_logos," layout-valign=\"bottom\"}")`]{.content-visible when-format="markdown"}

```{r}
#| label: print-logos
#| results: asis

# `cat` is used to avoid extra processing
# `sep` argument adds the line break and then a new line as required
if (params$outputFormat == "markdown") cat(logos$logo_image_line, sep="\n\n")
```

[`r paste(":::")`]{.content-visible when-format="markdown"}

`r build_banners(params$outputFormat, length(params$warningLevel$yellow) > 0, length(params$warningLevel$orange) > 0, length(params$warningLevel$red) > 0)`

The Regions of BC highlighted on the map are being impacted or are likely to be impacted by wildfire smoke over the next `r params$smokeDuration`.

::: {.content-visible when-format="pdf"}

\begin{wrapfigure}{r, t}{0.5\textwidth}
\centering
\vspace{-0.5cm}
\includegraphics[width= 0.55\textwidth]{`r map`}
\captionsetup{width = 0.45\textwidth, font = footnotesize, labelformat = empty, justification = raggedright}
\caption{Descriptions of the affected region(s) are provided at the end of this document. This Air Quality Warning excludes the area managed by Metro Vancouver. For air quality notifications issued by Metro Vancouver refer to Metro Vancouver's \href{https://metrovancouver.org/services/air-quality-climate-action/air-quality-data-and-advisories}{air quality data and warnings page}.}
\vspace*{-2 cm}
\end{wrapfigure}

:::

`r params$customMessage`

During a wildfire, smoke conditions can change quickly over short distances and can vary considerably hour-by-hour.

Monitor the situation for any wildfires near your area at [EmergencyInfoBC](https://www.emergencyinfobc.gov.bc.ca/). Follow all emergency announcements relevant to your location.

The next update will be available on `r format(as.Date(params$nextUpdate), '%B %d, %Y')` and made available on the Province's [Air Quality Warnings webpage](https://aqwarnings.gov.bc.ca/).

::: {.content-visible when-format="markdown"}

![Descriptions of affected region(s) are provided at the end of this page. This Air Quality Warning excludes the area managed by Metro Vancouver. Refer to the More Information section for a link to air quality notifications issued by Metro Vancouver.](`r map_html`){fig-alt="Air Quality Warning - Wildfire Smoke Regions for `r format(Sys.Date(), '%B %d, %Y')`: `r string_region_names`." width="100%" height="600px"}

:::

## Actions you can take
::: {.content-hidden when-meta="twoLevels"}

`r string_action_text`

`r string_all_levels`

:::

::: {.content-visible when-meta="twoLevels" when-format="markdown"}
  
  {{< accordion_controls >}}
  {{< accordion_start title="For regions under yellow warnings" initiallyOpen="true" variant="yellow">}}
  
:::
  
::: {.content-visible when-meta="twoLevels" when-format="pdf"}
  
 **For regions under yellow warnings**
  
:::
  
  `r string_yellow_text`
  
  `r string_all_levels`
  
::: {.content-visible when-meta="twoLevels" when-format="markdown"}
  
  {{< accordion_end >}}
  {{< accordion_start title="For regions under orange warnings" initiallyOpen="true" variant="orange">}}
  
:::
  
::: {.content-visible when-meta="twoLevels" when-format="pdf"}
  
  **For regions under orange warnings**
  
:::
  
  `r string_orange_text`
  
  `r string_all_levels`
  
  
::: {.content-visible when-meta="twoLevels" when-format="markdown"}
  
  {{< accordion_end >}}
  
:::

## During smoky conditions

::: {.content-visible when-format="markdown"}

{{< accordion_controls >}}
{{< accordion_start title="Follow your common sense" initiallyOpen="true" >}}

:::

::: {.content-visible when-format="pdf"}

### Follow your common sense

:::

-   Stop or reduce your activity level if breathing becomes uncomfortable or you feel unwell.

-   Stay cool and drink plenty of fluids.

-   Carry any rescue medications with you at all times.

-   Make sure that children and others who cannot care for themselves follow the same advice.

::: {.content-visible when-format="markdown"}

{{< accordion_end >}}
{{< accordion_start title="Monitor your symptoms" initiallyOpen="true" >}}

:::

::: {.content-visible when-format="pdf"}

### Monitor your symptoms

:::

-   Different people have different responses to smoke.

-   Mild irritation and discomfort such as eye, nose and throat irritation, phlegm production, runny nose, headaches or mild cough are common, and usually disappear when the smoke clears.

-   More serious but less common symptoms include wheezing, chest pains or severe cough.

-   People with asthma or other chronic illness should follow any personal care plans designed with their family physicians.

-   If you are unsure whether you need medical care, call HealthLink BC at 8-1-1.

-   If you are experiencing difficulty in breathing, chest pain or discomfort, or a severe cough, contact your physician, walk-in clinic, or emergency department. **If you are having a medical emergency, call 9-1-1.**

::: {.content-visible when-format="markdown"}

{{< accordion_end >}}
{{< accordion_start title="Tips to reduce your smoke exposure" initiallyOpen="true" >}}

:::

::: {.content-visible when-format="pdf"}

### Tips to reduce your smoke exposure

:::

-   Smoke levels may still be elevated indoors, so stay aware of your symptoms even when you are indoors.

-   When indoors, keep windows and doors closed. If it is hot, prioritize keeping cool.

-   Protect your indoor air from wildfire smoke. Use a clean, good quality air filter in your ventilation system and/or a certified portable air cleaner that can filter fine particles. Do-it-yourself air cleaners may also be used if other options are unavailable. For more details, see the BC Centre of Disease Control [Do-it Yourself Air Cleaners][DIY air cleaner fact sheet] fact sheet. 

-   If your home is smoky or hot, spend time in places with cleaner, air conditioned air such as community centres, libraries, or malls.

-   If you must be outside, a well-constructed, well-fitting and properly worn respirator type mask (such as a NIOSH-certified N95 or equivalent respirator) can reduce your exposure to the fine particles in the smoke. Even though exposure may be reduced, there can still be risks to health.

-   When in a vehicle, keep windows closed with the air flow (or air conditioning) set to recirculate.

-   Check on others who are in your care or live nearby who may be more likely to be impacted by wildfire smoke.

-   Always follow guidance from local authorities.

::: {.content-visible when-format="markdown"}

{{< accordion_end >}}

:::

## More information

::: {.content-visible when-format="markdown"}

**For additional general information about wildfire smoke and air quality:**

::: bcds-card-wrapper
  {{< card_start title="Air Quality Health Index" variant="info" logo="/assets/icon_air.svg" >}} Provincial summary of latest index and forecast.

  [What's the air like today?] {{< card_end >}} {{< card_start title="Air Quality Map" variant="info" logo="/assets/icon_air.svg" >}} Provincial air quality data refreshed every hour.

  [Latest air quality data] {{< card_end >}} {{< card_start title="Environment and Climate Change Canada FireWork" variant="danger" logo="/assets/icon_smoke.svg" >}} Air quality maps of next 72 hours.

  [Smoke forecast maps] {{< card_end >}} {{< card_start title="FireSmoke Canada" variant="danger" logo="/assets/icon_smoke.svg" >}} BlueSky Canada smoke forecasting system.

  [Smoke forecast map] {{< card_end >}} {{< card_start title="Metro Vancouver Air Quality" variant="success" logo="/assets/icon_chart.svg" >}} Warnings and map of air quality and weather data.

  [Air quality data and warnings] {{< card_end >}}
```{r, results='asis', echo=FALSE, eval= flag_run_pg}

  cat('{{< card_start title="Local Resources" variant="info" logo="/assets/icon_air.svg" >}}\n\n [Prince George Air Improvement Roundtable](https://www.pgairquality.com/) {{< card_end >}}\n\n')
```
:::
:::


::: {.content-visible when-format="pdf"}

### For additional information about wildfire smoke and air quality

Air quality notifications issued by Metro Vancouver

- Visit Metro Vancouver's [air quality data and warnings page][Air quality data and warnings]

Air quality data

- View the latest [Air Quality Health Index (AQHI) and forecast][What's the air like today?]

- Visit the provincial [air quality map][Latest air quality data] to view real-time air quality measurements

Wildfire smoke forecasts

- Environment and Climate Change Canada smoke forecasts - [FireWork][Smoke forecast maps]

- BlueSky Canada smoke forecasts - [FireSmoke Canada][Smoke forecast map]

```{r, results='asis', echo=FALSE, eval= flag_run_pg}
cat("Local resources\n\n")
cat("- Visit the [Prince George Air Improvement Roundtable's page](https://www.pgairquality.com/)\n\n")
```
:::

[Air quality data and warnings]: https://metrovancouver.org/services/air-quality-climate-action/air-quality-data-and-advisories
[What's the air like today?]: https://www.env.gov.bc.ca/epd/bcairquality/data/aqhi-table.html
[Latest air quality data]: https://www.env.gov.bc.ca/epd/bcairquality/readings/find-stations-map.html
[Smoke forecast maps]: https://weather.gc.ca/firework/index_e.html
[Smoke forecast map]: https://firesmoke.ca/forecasts/current/

::: {.content-visible when-format="markdown"}

**For additional general information about wildfire smoke and your health:**

<div class="bcds-card-wrapper">
{{< card_start title="BC Centre for Disease Control" logo="/assets/logo_BCCDC.png" >}} Fact sheets on wildfire smoke and its health impacts.

[Wildfire smoke] {{< card_end >}}`r HA_text`{{< card_start title="First Nations Health Authority" logo="/assets/logo_FNHA_horizontal.png" >}} Information and resources to help you prepare.

[Wildfire smoke health information] {{< card_end >}} {{< card_start title="HealthLink BC" logo="/assets/logo_BCID_HealthLinkBC_RGB_pos.png" >}} Learn and prepare for wildfires. Phone 8-1-1 for 24/7 health advice.

[Wildfires and your health] {{< card_end >}} {{< card_start title="Worksafe BC" logo="/assets/logo_worksafebc.jpg" >}} Info on working outdoors during smoky conditions.

[Wildfire FAQ] {{< card_end >}}
</div>

:::

::: {.content-visible when-format="pdf"}

### For additional general information about wildfire smoke and your health

BC Centre for Disease Control

- Wildfire smoke and its health impacts - [Fact sheets][Wildfire smoke]

Provincial Health Authorities

- First Nations Health Authority - [Wildfire smoke health information][Wildfire smoke health information]

`r HA_text`

HealthLink BC

- Learn about and prepare for wildfires - [Wildfires and your health resource][Wildfires and your health]

- Phone 8-1-1 for 24/7 health advice

Worksafe BC

- Working outdoors during smoky conditions - [Wildfires smoke FAQ][Wildfire FAQ]

:::

[Wildfire smoke]: http://www.bccdc.ca/health-info/prevention-public-health/wildfire-smoke
[DIY air cleaner Fact Sheet]: https://www.bccdc.ca/resource-gallery/Documents/Guidelines%20and%20Forms/Guidelines%20and%20Manuals/Health-Environment/BCCDC_WildFire_FactSheet_BoxFanAirFilters.pdf
[Wildfire smoke health information]: https://www.fnha.ca/about/news-and-events/news/wildfire-smoke-health-information
[Wildfires and your health]: https://www.healthlinkbc.ca/health-library/health-features/wildfires-and-your-health
[Wildfire FAQ]: https://www.worksafebc.com/en/resources/health-safety/information-sheets/wildfire-smoke-frequently-asked-questions-faq

## Contact

::: {.content-visible when-format="markdown"}

**Media questions regarding air quality conditions related to this Air Quality Warning:**
{{< card_start width="wide" >}}
:::

::: {.content-visible when-format="pdf"}

### Media questions regarding air quality conditions related to this Air Quality Warning
:::

`r get_meteorologist_contact(params$aqMet, markup_separator)`

::: {.content-visible when-format="markdown"}

{{< card_end >}}

**Media questions regarding health implications of wildfire smoke:**

{{< card_start  width="wide" >}}

:::

::: {.content-visible when-format="pdf"}

### Media questions regarding health implications of wildfire smoke

:::

`r paste(data_health_authority_contact$html_string, collapse = "\n\n")`[{{< card_end >}}]{.content-visible when-format="markdown"}

## Regions included under this Air Quality Warning

`r string_region_output`
